---
title: "GDSC CCLE common cell lines camera analysis"
output:
  html_document:
    df_print: paged
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE)
library(reshape)
library(GEOquery)
library(gplots)
library(dplyr)
library(GSVA)
library(GSEABase)
library(org.Hs.eg.db)
library(annotate)
library(GO.db)
library(limma)
library(reshape2)
library(biomaRt)
library(data.table)

```

```{r drugdata input 1}
ccle.expression<-fread('../data/CCLE/CCLE_Expression_Entrez_2012-09-29.gct',skip = 2, sep = '\t')
ccle.expression_o <- ccle.expression
ccle.expression <- as.data.frame(ccle.expression)
ccle.expression <- ccle.expression[,-c(1,2)]
ccle.anno <- ccle.expression_o[,c(1,2)]
toremove <- which(ccle.anno$Description == '')
ccle.anno <- ccle.anno[-toremove,]
ccle.expression<-ccle.expression[-toremove,] ## remove genes

load('../data/GDSC_CCLE_experiment_cl.Rdata')
load('../data/regulators_5_cl_20_tar.Rdata')
tokeep <- gdsc_ccle_exp$CCLE_NAME
reg.9 <- selected.regulator

drugdata<-read.csv('../data/CCLE/CCLE_NP24.2009_Drug_data_2015.02.24.csv')
drugdata.cast<-cast(drugdata,formula = CCLE.Cell.Line.Name ~ Compound, value = 'IC50..uM.')
row.names(drugdata.cast)<-drugdata.cast$CCLE.Cell.Line.Name
drugdata.cast$CCLE.Cell.Line.Name<-NULL
drugdata.cast<-as.data.frame(drugdata.cast)

regset <- getGmt('../data/Gene_regulators/targetset_nodup.gmt.txt')
drug.camera<-drugdata.cast[tokeep,]
ccle.camera<-ccle.expression[,tokeep]

ccle.drug.o <- drug.camera
ccle.tokeep <- ccle.camera


```

```{r drug}
drug.camera[drug.camera == 8] <- 'nr'
drug.camera[(! is.na(drug.camera)) & (drug.camera != 'nr')] <- 'r'
ccle.drug <- drug.camera
```


```{r camera 0 , echo = FALSE}
gene.ORF <- unlist(strsplit(ccle.anno$Name, "_"))
gene.ORF <- gene.ORF[gene.ORF != 'at']
gene.name <- ccle.anno$Description
ids <- ids2indices(geneIds(regset),gene.name) 
camera.drug<-function(x){
#x <- '17-AAG'
drug<-x
cl.drug<-drug.camera[,drug]
cl.name <- row.names(drug.camera)
cl.name <- cl.name[!is.na(cl.drug)]
cl.drug <- cl.drug[!is.na(cl.drug)]

n<-length(cl.drug)
cl.design <- matrix(nrow=n,ncol=2)
cl.design[,1] <- as.numeric(cl.drug == 'r')
cl.design.2<-cl.design[,1]
cl.design.2[which(cl.design[,1] == 1)]<- 0
cl.design.2[which(cl.design[,1] == 0)]<- 1
cl.design[,2]<-cl.design.2
colnames(cl.design)<-c('response','non_response')
r1 <- camera(data.matrix(ccle.camera[,cl.name]), ids, cl.design,contrast=makeContrasts(response - non_response, levels=cl.design))
#Pvalue<-r1$Pvalue
#Direction<-r1$Direction 
#Pvalue<-1-Pvalue
#Pvalue[which(Direction == 'Down')]<- -Pvalue[which(Direction == 'Down')]
return(r1[which(r1$PValue<0.05 & r1$NGenes > 9),])}

drugs <- colnames(drug.camera)
drugs <- drugs[! drugs == 'Panobinostat']
result.camera <- lapply(drugs,camera.drug)
names(result.camera) <- drugs

```

```{r common drug a}
drugs <- ccle.com.drug
```


```{r dge 0}
dge.drug<-function(x){
#x <- '17-AAG'
drug<-x

cl.drug<-drug.camera[,drug]
cl.name <- row.names(drug.camera)
cl.name <- cl.name[!is.na(cl.drug)]
cl.drug <- cl.drug[!is.na(cl.drug)]

n<-length(cl.drug)
cl.design <- matrix(nrow=n,ncol=2)
cl.design[,1] <- as.numeric(cl.drug == 'r')
cl.design.2<-cl.design[,1]
cl.design.2[which(cl.design[,1] == 1)]<- 0
cl.design.2[which(cl.design[,1] == 0)]<- 1
cl.design[,2]<-cl.design.2
colnames(cl.design)<-c('response','non_response')
fit <- lmFit(ccle.camera[which(gene.name %in% reg.9),cl.name], cl.design)
contrasts <- makeContrasts(response - non_response, levels=cl.design)
fit2 <- contrasts.fit(fit, contrasts)
fit2 <- eBayes(fit2)
fit2$genes <- ccle.anno[which(ccle.anno$Description %in% reg.9),]
topTable(fit2)
testResults <- topTable(fit2, number=nrow(fit2))
testResults <- testResults[complete.cases(testResults),]
return(testResults)
}

result.dge <- lapply(drugs,dge.drug)
names(result.dge) <- drugs
```


```{r camera go slim 0}
slim_terms <- getOBOCollection("../data/Gene_regulators/goslim_generic.obo")
xx <- as.list(org.Hs.egGO2ALLEGS)
goids <- xx[2:3]
gonames<-Term(GOTERM)
go_bp_children <- GOBPOFFSPRING$"GO:0008150"
slim_terms <- ids(slim_terms)[which(ids(slim_terms) %in% go_bp_children)]
slim_terms <- cbind(slim_terms, unname(gonames[match(slim_terms, names(gonames))]))
pathways <- xx[slim_terms[,1]]
pathways <- pathways[!is.na(names(pathways))]
names(pathways) <- unname(gonames[match(names(pathways), names(gonames))])
ids<-ids2indices(pathways,gene.ORF)
camera.drug.go<-function(x){
#x <- 'Panobinostat'
drug<-x
cl.drug<-drug.camera[,drug]
cl.name <- row.names(drug.camera)
cl.name <- cl.name[!is.na(cl.drug)]
cl.drug <- cl.drug[!is.na(cl.drug)]

n<-length(cl.drug)
cl.design <- matrix(nrow=n,ncol=2)
cl.design[,1] <- as.numeric(cl.drug == 'r')
cl.design.2<-cl.design[,1]
cl.design.2[which(cl.design[,1] == 1)]<- 0
cl.design.2[which(cl.design[,1] == 0)]<- 1
cl.design[,2]<-cl.design.2
colnames(cl.design)<-c('response','non_response')
r1 <- camera(data.matrix(ccle.camera[,cl.name]), ids, cl.design,contrast=makeContrasts(response - non_response, levels=cl.design))
#Pvalue<-r1$Pvalue
#Direction<-r1$Direction 
#Pvalue<-1-Pvalue
#Pvalue[which(Direction == 'Down')]<- -Pvalue[which(Direction == 'Down')]
return(r1[which(r1$PValue<0.05),])}


result.camera.goslim <- lapply(drugs,camera.drug.go)
names(result.camera.goslim) <- drugs
```

```{r camera go h 0}
hgset <- getGmt('../data/Gene_regulators/h.all.v6.2.symbols.gmt')
ids <- ids2indices(geneIds(hgset),gene.name) 
camera.drug.h<-function(x){
#x <- 'Panobinostat'
drug<-x
cl.drug<-drug.camera[,drug]
cl.name <- row.names(drug.camera)
cl.name <- cl.name[!is.na(cl.drug)]
cl.drug <- cl.drug[!is.na(cl.drug)]

n<-length(cl.drug)
cl.design <- matrix(nrow=n,ncol=2)
cl.design[,1] <- as.numeric(cl.drug == 'r')
cl.design.2<-cl.design[,1]
cl.design.2[which(cl.design[,1] == 1)]<- 0
cl.design.2[which(cl.design[,1] == 0)]<- 1
cl.design[,2]<-cl.design.2
colnames(cl.design)<-c('response','non_response')
r1 <- camera(data.matrix(ccle.camera[,cl.name]), ids, cl.design,contrast=makeContrasts(response - non_response, levels=cl.design))
#Pvalue<-r1$Pvalue
#Direction<-r1$Direction 
#Pvalue<-1-Pvalue
#Pvalue[which(Direction == 'Down')]<- -Pvalue[which(Direction == 'Down')]
return(r1[which(r1$PValue<0.05),])}

result.camera.h <- lapply(drugs,camera.drug.h)
names(result.camera.h) <- drugs
```

```{r camera go kegg 0}
keggset <- getGmt('../data/Gene_regulators/c2.cp.kegg.v6.2.symbols.gmt.txt')
ids <- ids2indices(geneIds(keggset),gene.name) 
camera.drug.kegg<-function(x){
#x <- 'Panobinostat'
drug<-x
cl.drug<-drug.camera[,drug]
cl.name <- row.names(drug.camera)
cl.name <- cl.name[!is.na(cl.drug)]
cl.drug <- cl.drug[!is.na(cl.drug)]

n<-length(cl.drug)
cl.design <- matrix(nrow=n,ncol=2)
cl.design[,1] <- as.numeric(cl.drug == 'r')
cl.design.2<-cl.design[,1]
cl.design.2[which(cl.design[,1] == 1)]<- 0
cl.design.2[which(cl.design[,1] == 0)]<- 1
cl.design[,2]<-cl.design.2
colnames(cl.design)<-c('response','non_response')
r1 <- camera(data.matrix(ccle.camera[,cl.name]), ids, cl.design,contrast=makeContrasts(response - non_response, levels=cl.design))
#Pvalue<-r1$Pvalue
#Direction<-r1$Direction 
#Pvalue<-1-Pvalue
#Pvalue[which(Direction == 'Down')]<- -Pvalue[which(Direction == 'Down')]
return(r1[which(r1$PValue<0.05),])}

result.camera.kegg <- lapply(drugs,camera.drug.kegg)
names(result.camera.kegg) <- drugs
```
## Example mek inhibitor AZD06244

### regulator target sets of differential expression by camera
```{r azd, }
result.camera$AZD6244
azd.name <- row.names(result.camera$AZD6244)
```

### differential expression of significant regulators by logFC

```{r azd2, }
azd <- result.dge$AZD6244
azd[which(azd$logFC > 1 | azd$logFC < -1),]
```

### differential expression of regulators

```{r dge2, }
azd[which(azd$genename %in% azd.name),]
```

### go slim
```{r goslim0, }
result.camera.goslim$AZD0530
```

### kegg slim
```{r keggslim0, }
result.camera.kegg$AZD0530
```

### hallmark 
```{r h0, }
result.camera.h$AZD0530
```

## Example mek inhibitor PD-0325901

### regulator target sets of differential expression by camera
```{r pd, }
result.camera$`PD-0325901`
pd.name <- row.names(result.camera$`PD-0325901`)
```

### differential expression of significant regulators by logFC

```{r pd2,  }
pd <- result.dge$`PD-0325901`
pd[which(pd$logFC > 1 | pd$logFC < -1),]
```

### differential expression of regulators
```{r dge3,  }
pd[which(pd$genename %in% pd.name),]
```

### go slim
```{r goslim, }
pd.go <- result.camera.goslim$`PD-0325901`
```

### kegg slim
```{r keggslim1, }
result.camera.kegg$`PD-0325901`
```

### hallmark 
```{r h1, }
result.camera.h$`PD-0325901`
```


```{r exp input}
gdsc.expression<-fread('../data/GDSC/sanger1018_brainarray_ensemblgene_rma.txt',sep = 'auto',check.names = FALSE,header = TRUE,blank.lines.skip=TRUE)
gdsc.expression<-as.data.frame(gdsc.expression)
row.names(gdsc.expression)<-gdsc.expression$ensembl_gene
gdsc.expression$ensembl_gene<-NULL
gdsc.exp <- gdsc.expression
tokeep <- gdsc_ccle_exp$COSMIC_ID
```

```{r drug input}
gdsc.drug<-read.csv("../data/GDSC/v17.3_fitted_dose_response.csv")
gdsc.tocast<-data.frame(gdsc.drug$COSMIC_ID,gdsc.drug$DRUG_NAME,gdsc.drug$LN_IC50)
colnames(gdsc.tocast)<-c('COSMIC_ID','drug','LN_IC50')
gdsc.cast<-dcast(gdsc.tocast,formula = COSMIC_ID ~ drug,fun.aggregate = mean)
row.names(gdsc.cast)<-gdsc.cast$COSMIC_ID
gdsc.cast$COSMIC_ID<-NULL
gdsc.cast <- exp(gdsc.cast)
gdsc.cast_o<-gdsc.cast
```

```{r drug convert}
gdsc.tomax<-data.frame(gdsc.drug$COSMIC_ID,gdsc.drug$DRUG_NAME,gdsc.drug$MAX_CONC_MICROMOLAR)
colnames(gdsc.tomax)<-c('COSMIC_ID','drug','MAX_CONC_MICROMOLAR')
gdsc.cast.max<-dcast(gdsc.tomax,formula = COSMIC_ID ~ drug,fun.aggregate = mean)
row.names(gdsc.cast.max)<-gdsc.cast.max$COSMIC_ID
gdsc.cast.max$COSMIC_ID <- NULL
cl.toresponse <- matrix( nrow = nrow(gdsc.cast), ncol = 251)
#regulator.drug.ssgsea.cor<-sapply(c(1:251),function(x){
  for (i in c(1:251)){
  dr <- gdsc.cast[,i]
  names(dr) <- row.names(gdsc.cast)
  cl.thres <- gdsc.cast.max[names(dr),i]
  dr[dr < cl.thres] <- 'r'
  dr[!(dr == 'NaN' | dr == 'r')] <- 'nr'
  #dr[dr >= cl.thres] <- 'nr'
  cl.toresponse[,i] <- dr
  }
cl.toresponse <- as.data.frame(cl.toresponse)
row.names(cl.toresponse) <- row.names(gdsc.cast)
colnames(cl.toresponse) <- colnames(gdsc.cast)
```



```{r load}
#load('../data/GDSC/gdsc_exp_nohemo_z.Rdata')
```


```{r info}
genes <- row.names(gdsc.exp)
ensembl <- useDataset("hsapiens_gene_ensembl", useMart("ensembl"))
gene_list <- getBM(filters= "ensembl_gene_id", attributes= c("ensembl_gene_id",
"entrezgene", 'hgnc_symbol',"description"),values=genes,mart= ensembl)
gene.name <- gene_list$hgnc_symbol
gene.ORF <- gene_list$entrezgene
gene.list <- gene_list$ensembl_gene_id
```

```{r pre}
gdsc.camera<-gdsc.exp[match(gene.list,genes),tokeep]

drug.camera<-cl.toresponse[tokeep,]

gdsc.drug <- drug.camera

drug.tokeep <- gdsc.cast_o[tokeep,]
```



```{r camera ,  }
ids <- ids2indices(geneIds(regset),gene.name) 
camera.drug<-function(x){
#x <- 'Erlotinib'
drug<-x
cl.drug<-drug.camera[,drug]
names(cl.drug)<-row.names(drug.camera)
cl.name<-row.names(drug.camera)
cl.drug <- cl.drug[!(cl.drug == 'NaN')]
n<-length(cl.drug)
cl.design <- matrix(nrow=n,ncol=2)
cl.design[,1] <- as.numeric(cl.drug == 'r')
cl.design.2<-cl.design[,1]
cl.design.2[which(cl.design[,1] == 1)]<- 0
cl.design.2[which(cl.design[,1] == 0)]<- 1
cl.design[,2]<-cl.design.2
colnames(cl.design)<-c('response','non_response')
r1 <- camera(data.matrix(gdsc.camera[,names(cl.drug)]), ids, cl.design,contrast=makeContrasts(response - non_response, levels=cl.design))
#Pvalue<-r1$Pvalue
#Direction<-r1$Direction 
#Pvalue<-1-Pvalue
#Pvalue[which(Direction == 'Down')]<- -Pvalue[which(Direction == 'Down')]
return(r1[which(r1$PValue<0.05 & r1$NGenes > 9),])}
#r1[which(r1$PValue<0.05 & r1$NGenes > 9),]
drugs <- colnames(drug.camera)
#drugs <- drugs[! drugs == 'Panobinostat']
mek_in <- c('Refametinib','CI-1040','PD0325901','Selumetinib','Trametinib')
result.camera <- lapply(mek_in,camera.drug)
names(result.camera) <- mek_in
#camera.drug('Erlotinib')
```

```{r, eval = FALSE}
common.drug[common.drug %in% ccle.drug]
common.drug[! common.drug%in% ccle.drug]
ccle.com.drug <- common.drug
```

```{r, eval = FALSE}
common.drug
```

```{r, eval = FALSE}
ccle.com.drug[7] <- '17-AAG'
ccle.com.drug[8] <- 'AZD0530'
ccle.com.drug[9] <-  "AZD6244"
ccle.com.drug[10] <- 'TAE684'
ccle.com.drug[11] <- "PLX4720"
ccle.com.drug[12] <- "Nutlin-3"  
ccle.com.drug[13] <- "PD-0325901"
ccle.com.drug[14] <- "PD-0332991"
```

```{r}
ccle.com.drug
```

```{r dge}
gdsc.expression<-fread('../data/GDSC/sanger1018_brainarray_ensemblgene_rma.txt',sep = 'auto',check.names = FALSE,header = TRUE,blank.lines.skip=TRUE)
gdsc.expression<-as.data.frame(gdsc.expression)
row.names(gdsc.expression)<-gdsc.expression$ensembl_gene
gdsc.expression$ensembl_gene<-NULL
dge.drug<-function(x){
#x <- '17-AAG'
drug<-x
cl.drug<-drug.camera[,drug]
names(cl.drug)<-row.names(drug.camera)
cl.name<-row.names(drug.camera)
cl.drug <- cl.drug[!(cl.drug == 'NaN')]
n<-length(cl.drug)
cl.design <- matrix(nrow=n,ncol=2)
cl.design[,1] <- as.numeric(cl.drug == 'r')
cl.design.2<-cl.design[,1]
cl.design.2[which(cl.design[,1] == 1)]<- 0
cl.design.2[which(cl.design[,1] == 0)]<- 1
cl.design[,2]<-cl.design.2
colnames(cl.design)<-c('response','non_response')
fit <- lmFit(gdsc.expression[which(gene.name %in% reg.9),names(cl.drug)], cl.design)
contrasts <- makeContrasts(response - non_response, levels=cl.design)
fit2 <- contrasts.fit(fit, contrasts)
fit2 <- eBayes(fit2)
fit2$genes <- gene_list[which(gene.name %in% reg.9),]
topTable(fit2)
testResults <- topTable(fit2, number=nrow(fit2))
testResults <- testResults[complete.cases(testResults),]
return(testResults)
}

result.dge.gd <- lapply(common.drug,dge.drug)
names(result.dge.gd) <- common.drug
```

```{r camera go slim, }
slim_terms <- getOBOCollection("../data/Gene_regulators/goslim_generic.obo")
xx <- as.list(org.Hs.egGO2ALLEGS)
goids <- xx[2:3]
gonames<-Term(GOTERM)
go_bp_children <- GOBPOFFSPRING$"GO:0008150"
slim_terms <- ids(slim_terms)[which(ids(slim_terms) %in% go_bp_children)]
slim_terms <- cbind(slim_terms, unname(gonames[match(slim_terms, names(gonames))]))
pathways <- xx[slim_terms[,1]]
pathways <- pathways[!is.na(names(pathways))]
names(pathways) <- unname(gonames[match(names(pathways), names(gonames))])
ids<-ids2indices(pathways,gene.ORF)
camera.drug.go<-function(x){
#x <- 'Panobinostat'
drug<-x
cl.drug<-drug.camera[,drug]
names(cl.drug)<-row.names(drug.camera)
cl.name<-row.names(drug.camera)
cl.drug <- cl.drug[!(cl.drug == 'NaN')]
n<-length(cl.drug)
cl.design <- matrix(nrow=n,ncol=2)
cl.design[,1] <- as.numeric(cl.drug == 'r')
cl.design.2<-cl.design[,1]
cl.design.2[which(cl.design[,1] == 1)]<- 0
cl.design.2[which(cl.design[,1] == 0)]<- 1
cl.design[,2]<-cl.design.2
colnames(cl.design)<-c('response','non_response')
r1 <- camera(data.matrix(gdsc.camera[,names(cl.drug)]), ids, cl.design,contrast=makeContrasts(response - non_response, levels=cl.design))
#Pvalue<-r1$Pvalue
#Direction<-r1$Direction 
#Pvalue<-1-Pvalue
#Pvalue[which(Direction == 'Down')]<- -Pvalue[which(Direction == 'Down')]
return(r1[which(r1$PValue<0.05),])}

result.go.gd <- lapply(common.drug,camera.drug.go)
names(result.go.gd) <- common.drug

```

```{r camera go h, }
hgset <- getGmt('../data/Gene_regulators/h.all.v6.2.symbols.gmt')
ids <- ids2indices(geneIds(hgset),gene.name) 
camera.drug.h<-function(x){
#x <- 'Panobinostat'
#x <- 'Panobinostat'
drug<-x
cl.drug<-drug.camera[,drug]
names(cl.drug)<-row.names(drug.camera)
cl.name<-row.names(drug.camera)
cl.drug <- cl.drug[!(cl.drug == 'NaN')]
n<-length(cl.drug)
cl.design <- matrix(nrow=n,ncol=2)
cl.design[,1] <- as.numeric(cl.drug == 'r')
cl.design.2<-cl.design[,1]
cl.design.2[which(cl.design[,1] == 1)]<- 0
cl.design.2[which(cl.design[,1] == 0)]<- 1
cl.design[,2]<-cl.design.2
colnames(cl.design)<-c('response','non_response')
r1 <- camera(data.matrix(gdsc.camera[,names(cl.drug)]), ids, cl.design,contrast=makeContrasts(response - non_response, levels=cl.design))
#Pvalue<-r1$Pvalue
#Direction<-r1$Direction 
#Pvalue<-1-Pvalue
#Pvalue[which(Direction == 'Down')]<- -Pvalue[which(Direction == 'Down')]
return(r1[which(r1$PValue<0.05),])}

result.camera.h.gd <- lapply(common.drug,camera.drug.h)
names(result.camera.h.gd) <- common.drug
```

```{r camera go kegg,  }
keggset <- getGmt('../data/Gene_regulators/c2.cp.kegg.v6.2.symbols.gmt.txt')
ids <- ids2indices(geneIds(keggset),gene.name) 
camera.drug.kegg<-function(x){
#x <- 'Panobinostat'
#x <- 'Panobinostat'
drug<-x
cl.drug<-drug.camera[,drug]
names(cl.drug)<-row.names(drug.camera)
cl.name<-row.names(drug.camera)
cl.drug <- cl.drug[!(cl.drug == 'NaN')]
n<-length(cl.drug)
cl.design <- matrix(nrow=n,ncol=2)
cl.design[,1] <- as.numeric(cl.drug == 'r')
cl.design.2<-cl.design[,1]
cl.design.2[which(cl.design[,1] == 1)]<- 0
cl.design.2[which(cl.design[,1] == 0)]<- 1
cl.design[,2]<-cl.design.2
colnames(cl.design)<-c('response','non_response')
r1 <- camera(data.matrix(gdsc.camera[,names(cl.drug)]), ids, cl.design,contrast=makeContrasts(response - non_response, levels=cl.design))
#Pvalue<-r1$Pvalue
#Direction<-r1$Direction 
#Pvalue<-1-Pvalue
#Pvalue[which(Direction == 'Down')]<- -Pvalue[which(Direction == 'Down')]
return(r1[which(r1$PValue<0.05),])}

result.camera.kegg.gd <- lapply(common.drug,camera.drug.kegg)
names(result.camera.kegg.gd) <- common.drug
```

## Selumetinib (AZD6244) 
```{r AZD,eval = false}
result.camera$Selumetinib
se<-result.dge.gd$Selumetinib
se[which(se$logFC > 1 | se$logFC < -1),]
result.camera.h.gd$Selumetinib
result.camera.kegg.gd$Selumetinib
result.go.gd$Selumetinib
```

## PD0325901
```{r pf,eval=false}
result.camera$PD0325901
pd<-result.dge.gd$PD0325901
pd[which(pd$logFC > 1 | pd$logFC < -1),]
result.camera.h.gd$PD0325901
result.camera.kegg.gd$PD0325901
pd.go.g <- result.go.gd$PD0325901
```
```{r gene correlation}
cl<-colnames(ccle.camera)
setkey(gdsc_ccle_exp,'CCLE_NAME')
cosmic<-gdsc_ccle_exp[cl]$COSMIC_ID
gdsc.cosmic<-gdsc.camera[,cosmic]
gene.name.ccle <- ccle.anno$Description
gene.name.gdsc <- gene.name
ccle.camera.gene <- ccle.camera[match(gene.name.gdsc,gene.name.ccle),]
gene.corr <- c()
row.gene <- row.names(ccle.camera.gene)
for (i in c(1:17555)){
if (!grepl('NA', row.gene[i]))
{gene.corr[i]<- cor.test(as.numeric(ccle.camera.gene[i,]),as.numeric(gdsc.cosmic[i,]),method = 'pearson')$estimate}
else(gene.corr[i]<-NA)
}
```


```{r common response cell line}
result.cor.drug <- list()
for(i in c(1:length(common.drug))){

drug.gdsc <- common.drug[i]
drug.ccle <- ccle.com.drug[i]

gdsc.r <- which(as.character(gdsc.drug[,drug.gdsc]) == 'r')
ccle.r <- which(as.character(ccle.drug[,drug.ccle]) == 'r')

ccle.com.name <- gdsc_ccle_exp$CCLE_NAME

tokeep.r.g <- row.names(gdsc.drug)[gdsc.r]

tokeep.r.c <- row.names(ccle.drug)[ccle.r]

setkey(gdsc_ccle_exp,'COSMIC_ID')

tocor.g.c <- gdsc_ccle_exp[tokeep.r.g]$CCLE_NAME

comm.r.c <- intersect(tocor.g.c,tokeep.r.c)

if (length(comm.r.c) >= 3){
setkey(gdsc_ccle_exp,'CCLE_NAME')

comm.r.g <-  gdsc_ccle_exp[comm.r.c]$COSMIC_ID

result.cor.drug[[i]] <- cor.test(drug.tokeep[comm.r.g,drug.gdsc],ccle.drug.o[comm.r.c,drug.ccle])
}else{result.cor.drug[[i]] <- NA}
}

names(result.cor.drug) <- common.drug
result.cor.drug
#}
```
```{r}
i <- 13
drug.gdsc <- common.drug[i]
drug.ccle <- ccle.com.drug[i]

gdsc.r <- which(as.character(gdsc.drug[,drug.gdsc]) == 'r')
ccle.r <- which(as.character(ccle.drug[,drug.ccle]) == 'r')

ccle.com.name <- gdsc_ccle_exp$CCLE_NAME

tokeep.r.g <- row.names(gdsc.drug)[gdsc.r]

tokeep.r.c <- row.names(ccle.drug)[ccle.r]

setkey(gdsc_ccle_exp,'COSMIC_ID')

tocor.g.c <- gdsc_ccle_exp[tokeep.r.g]$CCLE_NAME

comm.r.c <- intersect(tocor.g.c,tokeep.r.c)

setkey(gdsc_ccle_exp,'CCLE_NAME')

comm.r.g <-  gdsc_ccle_exp[comm.r.c]$COSMIC_ID

pd.skin <- comm.r.c[grep('SKIN',comm.r.c)]
```

```{r}

i <- 9
drug.gdsc <- common.drug[i]
drug.ccle <- ccle.com.drug[i]

gdsc.r <- which(as.character(gdsc.drug[,drug.gdsc]) == 'r')
ccle.r <- which(as.character(ccle.drug[,drug.ccle]) == 'r')

ccle.com.name <- gdsc_ccle_exp$CCLE_NAME

tokeep.r.g <- row.names(gdsc.drug)[gdsc.r]

tokeep.r.c <- row.names(ccle.drug)[ccle.r]

setkey(gdsc_ccle_exp,'COSMIC_ID')

tocor.g.c <- gdsc_ccle_exp[tokeep.r.g]$CCLE_NAME

comm.r.c <- intersect(tocor.g.c,tokeep.r.c)

setkey(gdsc_ccle_exp,'CCLE_NAME')

comm.r.g <-  gdsc_ccle_exp[comm.r.c]$COSMIC_ID

azd.skin <- comm.r.c[grep('SKIN',comm.r.c)]

```





```{r azd lung cell line}
cl.name<-row.names(ccle.drug)
gdsc.azd <- gdsc.drug$Selumetinib
gdsc.cos <- row.names(gdsc.drug) 

names(gdsc.azd) <- cl.name

gdsc.response <- cl.name[gdsc.azd == 'r']
ccle.azd <- ccle.drug$AZD6244
ccle.response <- cl.name[ccle.azd == 'r']
common.azd.r <- intersect(gdsc.response,ccle.response)
gdsc.nr <- cl.name[gdsc.azd == 'nr'] 
ccle.nr <- cl.name[ccle.azd == 'nr']
common.azd.nr <- intersect(gdsc.nr,ccle.nr)
common.azd.nr.lung <- common.azd.nr[grep('LUNG',common.azd.nr)]
common.azd.r.lung <- common.azd.r[grep('LUNG',common.azd.r)]
lung.common.cl <- c(common.azd.r.lung,common.azd.nr.lung)
save(lung.common.cl,file = '../data/lung_common_cl.Rdata')
```
```{r pd lung cell line}
gdsc.pd <- gdsc.drug$PD0325901
gdsc.response <- cl.name[gdsc.pd == 'r']
ccle.pd <- ccle.drug$`PD-0325901`
ccle.response <- cl.name[ccle.pd == 'r']
common.pd.r <- intersect(gdsc.response,ccle.response)
gdsc.nr <- cl.name[gdsc.pd == 'nr'] 
ccle.nr <- cl.name[ccle.pd == 'nr']
common.pd.nr <- intersect(gdsc.nr,ccle.nr)
common.pd.nr.lung <- common.pd.nr[grep('LUNG',common.pd.nr)]
common.pd.r.lung <- common.pd.r[grep('LUNG',common.pd.r)]
lung.common.cl <- c(common.pd.r.lung,common.pd.nr.lung)
save(lung.common.cl,file = '../data/lung_common_cl_pd.Rdata')
```

```{r skin cell line}
common.azd.nr.skin <- common.azd.nr[grep('SKIN',common.azd.nr)]
common.azd.r.skin <- common.azd.r[grep('SKIN',common.azd.r)]
skin.common.cl <- c(common.azd.r.skin,common.azd.nr.skin)
skin.cl <- cl.name[grep('SKIN',cl.name)]
gdsc.skin.azd <- gdsc.azd[skin.cl]
gdsc.skin.azd[ !  skin.cl  %in% skin.common.cl]
```

### common respnse drug correlation (needs revision)
```{r common un response cell line}
cl.com.ccle<-temp.ccle[gdsc.drug$Selumetinib[which(gdsc.drug$Selumetinib == ccle.drug$AZD6244)] == 'nr']
ccle.azd <- ccle.drug.o[cl.com.ccle,'AZD6244']
names(ccle.azd) <- cl.com.ccle
ccle.azd.ordered <- ccle.azd[order(ccle.azd)]

setkey(gdsc_ccle_exp,'CCLE_NAME')
ccle.azd.tb<-gdsc_ccle_exp[names(ccle.azd.ordered)]
ccle.azd.tb$dr <- ccle.azd.ordered
ccle.azd.tb

temp.gdsc <-row.names(gdsc.drug)[which(gdsc.drug$Selumetinib == ccle.drug$AZD6244)] 
cl.com.gdsc <- temp.gdsc[gdsc.drug$Selumetinib[which(gdsc.drug$Selumetinib == ccle.drug$AZD6244)] == 'nr']
gdsc.azd <- drug.tokeep[cl.com.gdsc,'Selumetinib']
names(gdsc.azd) <- cl.com.gdsc

gdsc.azd.ordered <- gdsc.azd[order(gdsc.azd)]

setkey(gdsc_ccle_exp,'COSMIC_ID')
gdsc.azd.tb<-gdsc_ccle_exp[names(gdsc.azd.ordered)]
gdsc.azd.tb$dr <- gdsc.azd.ordered
gdsc.azd.tb
```



```{r azd cor}
cor.test(ccle.azd,gdsc.azd,method = 'spearman')
```
```{r pd com}
temp.ccle <-row.names(ccle.drug)[which(gdsc.drug$PD0325901 == ccle.drug$`PD-0325901`)]
pd <- temp.ccle[gdsc.drug$PD0325901[which(gdsc.drug$PD0325901 == ccle.drug$`PD-0325901`)] == 'nr']
cl.com.ccle<-temp.ccle[gdsc.drug$PD0325901[which(gdsc.drug$PD0325901 == ccle.drug$`PD-0325901`)] == 'r']
ccle.pd <- ccle.drug.o[cl.com.ccle,'PD-0325901']

names(ccle.pd) <- cl.com.ccle
ccle.pd.ordered <- ccle.pd[order(ccle.pd)]

setkey(gdsc_ccle_exp,'CCLE_NAME')
ccle.pd.tb<-gdsc_ccle_exp[names(ccle.pd.ordered)]
ccle.pd.tb$dr <- ccle.pd.ordered
ccle.pd.tb


temp.gdsc <-row.names(gdsc.drug)[which(gdsc.drug$PD0325901 == ccle.drug$`PD-0325901`)]
cl.com.gdsc <- temp.gdsc[gdsc.drug$PD0325901[which(gdsc.drug$PD0325901 == ccle.drug$`PD-0325901`)] == 'r']
gdsc.pd <- drug.tokeep[cl.com.gdsc,'PD0325901']

names(gdsc.pd) <- cl.com.gdsc

gdsc.pd.ordered <- gdsc.pd[order(gdsc.pd)]

setkey(gdsc_ccle_exp,'COSMIC_ID')
gdsc.pd.tb<-gdsc_ccle_exp[names(gdsc.pd.ordered)]
gdsc.pd.tb$dr <- gdsc.pd.ordered
gdsc.pd.tb



```
```{r pd cor}
cor.test(ccle.pd,gdsc.pd,method = 'spearman')
```

### common non-response drug
```{r coss}
pd
```


### SLCO1B3 expression correlation

```{r response cl}

slco.cl <- gdsc_ccle_exp$CCLE_NAME
slco.cos <- gdsc_ccle_exp$COSMIC_ID
slco.doc <- gdsc.drug[slco.cos,'Docetaxel']
names(slco.doc) <- slco.cl

dc.r.name<-slco.cos[which(slco.doc == 'r')]

dc.drug<-drug.tokeep[dc.r.name,'Docetaxel']


gc.ccle.exp <- ccle.expression[which(ccle.anno$Description == 'SLCO1B3'),slco.cl[which(slco.doc == 'r')]]

cor.test(dc.drug,as.numeric(gc.ccle.exp[1,]),method = 'spearman')
dc.drug
as.numeric(gc.ccle.exp[1,])
```

```{r response cl prostate}
pro <- gdsc_ccle_exp[Site == 'prostate']
pro.cl <- pro$CCLE_NAME
pro.cos <- pro$COSMIC_ID
pro.doc <- gdsc.drug[pro.cos,'Docetaxel']
names(pro.doc) <- pro.cl

dc.r.name<-pro.cos[which(pro.doc == 'r')]

dc.drug<-drug.tokeep[dc.r.name,'Docetaxel']


gc.ccle.exp <- ccle.expression[which(ccle.anno$Description == 'SLCO1B3'),pro.cl[which(pro.doc == 'r')]]

cor.test(dc.drug,as.numeric(gc.ccle.exp[1,]),method = 'spearman')
dc.drug
as.numeric(gc.ccle.exp[1,])
```



```{r slco}
slco.exp<-ccle.tokeep[which(ccle.anno$Description == 'SLCO1B3'),]
doc.exp <- drug.tokeep[,'Docetaxel']
cor.test(as.numeric(doc.exp),as.numeric(slco.exp),method = 'spearman')
as.numeric(doc.exp)
as.numeric(slco.exp)
```



