---
title: "Copy number alteration analysis"
author: "Bingru Sun"
date: "2018/11/29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warnings = FALSE)
library(GSEABase)
```

## Method

compute copy number alteration change ratio by 1) extracting cell lines with mutated regulators, and the cell lines with non-mutated regulator. 2) count the cnv change by setting cutoff based on the data in all cell lines and all genes 3) compute the ratio of cnv change that is higher in not mutated cells lines of a regulator

```{r regulator mutation input,include = FALSE}
load('./data/selected_cl.Rdata')

mutation_o<-read.delim('./data/CCLE/CCLE_DepMap_18Q1_maf_20180207 (1).txt')
mutation<-mutation_o
mutation_s<-mutation_o[which(is.na(mutation_o$ExAC_AF)),]
mutation_s<-mutation_s[which(mutation_s$isDeleterious == 'TRUE'),]
mutation_s_w<-mutation_s
mutation_s_w<-data.frame(mutation_s$Hugo_Symbol,mutation_s$Tumor_Sample_Barcode)

cl.mut <- unique(mutation$Tumor_Sample_Barcode)

genelist<-as.character(unique(mutation_s_w[,1]))
rs<-function(x){
  c<-as.character(mutation_s_w[which(mutation_s_w$mutation_s.Hugo_Symbol==x),2])
  return(c)
  }
a<-lapply(genelist,rs)
names(a)<-genelist
mutatedcelllines<-a
b<-lapply(mutatedcelllines,length)

regset <- getGmt('./data/Gene_regulators/targetset_nodup.gmt.txt')
reg.list<-geneIds(regset)
reg.count<-lapply(reg.list,length)
reg.count<-unlist(reg.count)
  
regulator<-names(reg.count)
target <- unique(unlist(reg.list))

reg_m<-matrix(nrow=length(regulator),ncol=length(target))
colnames(reg_m)<-target
row.names(reg_m)<-regulator


for (i in 1:nrow(reg_m)){
  x <- regulator[i]
  for (y in reg.list[[i]])
  {reg_m[x,y] <- 1}
}

reg_m[is.na(reg_m)]<-0

```

## Distribution of CNA change ratio of all regulators

```{r info}

CNV_cell_lines_o <- read.table('./data/CCLE/CCLE_copynumber_byGene_2013-12-03.txt', header=TRUE, sep="\t")
CNV_cell_lines <- CNV_cell_lines_o
CNV_cell_lines$EGID <- NULL
CNV_cell_lines$CHR <- NULL
CNV_cell_lines$CHRLOC <- NULL
CNV_cell_lines$CHRLOCEND <- NULL

cl.cnv <- cl.tokeep

library(reshape2)
CNV_cell_lines.m <- melt(CNV_cell_lines)
colnames(CNV_cell_lines.m) <- c("Regulator", "Cell_line", "CNV")
CNV_cell_lines_m <- CNV_cell_lines.m[which(CNV_cell_lines.m$Cell_line %in% cl.cnv),]


mean(CNV_cell_lines_m$CNV)
summary(CNV_cell_lines_m$CNV)
plot(density(CNV_cell_lines_m$CNV),main = 'Distribution of CNA values in all cell lines')
#del_cul was 0.1 and amp_cut was 0.9
del_cut <- quantile(CNV_cell_lines_m$CNV[CNV_cell_lines_m$CNV < 0], 0.1)
amp_cut <- quantile(CNV_cell_lines_m$CNV[CNV_cell_lines_m$CNV > 0], 0.9)

#in mutated cell line 1, not in mutated cell line 0
choose<-function(x){return(ifelse(cl.cnv%in%mutatedcelllines[[x]],1,0))}
m<-sapply(regulator,choose)
m<-t(m)
colnames(m)<-cl.cnv
row.names(m)<-regulator

#########################################################
m_n<-m[,cl.cnv]
row.names(CNV_cell_lines)<-CNV_cell_lines$SYMBOL
extract_targetcl<-function(x){
  cl_x<-cl.cnv[which(m_n[x,]==1)]
  target_x<-target[which(reg_m[x,]==1)]
  result <- as.data.frame(CNV_cell_lines[target_x,cl_x])
  return(result)
}
findcnvchange<-function(x){sum(x > amp_cut | x < del_cut,na.rm =TRUE)}
findcnvchange_per<-function(x){sum(x > amp_cut | x < del_cut,na.rm =TRUE)/length(x)}
extract_target_notmutatedcl<-function(x){
  cl_x<-cl.cnv[which(m_n[x,]==0)]
  target_x<-target[which(reg_m[x,]==1)]
  result <- as.data.frame(CNV_cell_lines[target_x,cl_x])
  return(result)
}
############################################
computemutatedcnv<-function(x){
t<-extract_targetcl(x)
t_cnv<-apply(t,1,findcnvchange_per)
return(t_cnv)
}
computenotmutatedcnv<-function(x){
t<-extract_target_notmutatedcl(x)
t_cnv<-apply(t,1,findcnvchange_per)
return(t_cnv)
}
computecnv_diff<-function(x){
t<-extract_targetcl(x)
t_cnv<-apply(t,1,findcnvchange_per)
t_nm<-extract_target_notmutatedcl(x)
t_nm_cnv<-apply(t_nm,1,findcnvchange_per)
return(sum(t_nm_cnv>t_cnv)/nrow(t))
}
temp<-c()
n<-1
for (i in regulator){
  temp[n]<-computecnv_diff(i)
  n<-n+1
}

names(temp)<-regulator
cnvchange<-temp

a<-density(cnvchange,na.rm =TRUE)

plot(a,main = 'Distribution of the CNA change ratio per regulator')
```


```{r analysis}
reg_tar<-rowSums(reg_m)

reg_tar.sel<-regulator[reg_tar>= 20]

cnvchange<-cnvchange[!is.na(cnvchange)]

###############################

mutation_df <- sapply(regulator,function(x){return( cl.cnv %in% mutatedcelllines[[x]])})

row.names(mutation_df) <- cl.cnv

mutation_df <- t(mutation_df)

mt.count <- rowSums(mutation_df)
#############################

regulator.5<- regulator[mt.count >= 5]

selected.regulator <- intersect(regulator.5,reg_tar.sel)

cnvchange.5.tar <- cnvchange[ selected.regulator]

```

## Distribution of CNA change ratio of selecter regulators
* selected regulators (mutated in at least 5 cell lines and at least 20 target genes)

```{r plot}
plot(density(cnvchange.5.tar),main = 'distribution of CNA change ratio of selected regulators')

summary(cnvchange.5.tar)

cnv.genes<-names(cnvchange.5.tar[cnvchange.5.tar>0.5])
```

## regulators of ratio >= 0.5
```{r cna}
cnv.genes
```

## histogram and density plot of mutated cell line distribution

```{r hist}
summary(mt.count)
hist(mt.count,main = 'Number of mutated cell lines per regulators',xlab = 'number of cell lines')
```

```{r wilcox}
result<-c()

CNV_cell_lines_m <- data.table::data.table(CNV_cell_lines_m,stringsAsFactors = FALSE)

CNV_cell_lines_m[,Regulator := as.character(Regulator)]

CNV_cell_lines_m[,Cell_line := as.character(Cell_line)]

selected.regulator <- intersect(regulator.5,reg_tar.sel)

wilcox.regulator <- function(x){

target_x <- target[which(reg_m[x,]==1)]

cl_not_mutated <- cl.cnv[m_n[x,]==0]

cl_mutated <- cl.cnv[m_n[x,] == 1]

target_x<-target[which(reg_m[x,]==1)]

towilcox <- CNV_cell_lines_m[Regulator %in% target_x]

not_mutated_cl_cnv <- towilcox[Cell_line %in% cl_not_mutated]$CNV

mutated_cl_cnv <- towilcox[Cell_line %in% cl_mutated]$CNV

a <- wilcox.test(abs(mutated_cl_cnv),abs(not_mutated_cl_cnv),alternative = 'less')

return(a$p.value)

}

result.wilcox <- sapply(selected.regulator,wilcox.regulator)
```

## selected regulators in wilcoxon tests (p-value <0.05)
```{r wilcox result}
selected.regulator[result.wilcox <= 0.05]
wilcox.result <- selected.regulator[result.wilcox <= 0.05]
```

## selected regulators in both analysis (wilcoxon tests p-value <0.05, change ratio >= 0.5)
```{r regulators}
cnv.regulator <- intersect(cnv.genes,wilcox.result)
cnv.regulator
```

```{r save result, eval = FALSE}
save(selected.regulator, file = '../data/regulators_5_cl_20_tar.Rdata')
```

