---
title: "Drug correlation"
output:
  html_document: default
  html_notebook: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE)
library(reshape)
library(GEOquery)
library(gplots)
library(dplyr)
library(heatmaply)
library(GSVA)
library(GSEABase)
library(knitr)
library(kableExtra)
library(org.Hs.eg.db)
library(annotate)
```

# drug correlation analysis with target sets
```{r gettargetexpress,eval = FALSE}
mydata<-getGEO(filename='../data/GSE36133_series_matrix.txt')
pd<-pData(mydata)
toremove<-as.character(pd$`primary site:ch1`)
toremove<-toremove == 'haematopoietic_and_lymphoid_tissue'
gene.expression<-exprs(mydata)
gene.expression.nohemo<-gene.expression[,-which(toremove == TRUE)]
gene.expression.nohemo.z <- apply(gene.expression.nohemo,2,scale)
row.names(gene.expression.nohemo.z)<-rownames(gene.expression.nohemo)
#gene.expression.nohemo.z<-t(gene.expression.nohemo.z)
write.csv(gene.expression.nohemo.z,file = '../data/CCLE/geneexpression_zscore.csv')
```

```{r ssgsea,eval=FALSE}
gene.expression.nohemo.z<-read.csv('../data/CCLE/geneexpression_zscore.csv',row.names = 1)
mydata<-getGEO(filename='../data/GSE36133_series_matrix.txt')
#regset <- getGmt('../data/Gene_regulators/targetset.gmt.txt')
regset <- getGmt('../data/Gene_regulators/targetset_nodup.gmt.txt')
anno <- fData(mydata)
library(org.Hs.eg.db)
library(annotate)
ent<-as.character(anno$ORF)
genename<-getSYMBOL(ent,data='org.Hs.eg')
anno <- cbind(anno,genename)
gene.expression.nohemo.z<-gene.expression.nohemo.z[complete.cases(anno),]
anno<-anno[complete.cases(anno),]
rownames(gene.expression.nohemo.z)<-anno$genename
regsva<-gsva(data.matrix(gene.expression.nohemo.z),regset,method=c('ssgsea'))
save(regsva,file ='../data/regsva_new_z_col.Rdata')

```
```{r input,echo = FALSE}
load('/Users/ll/Desktop/git/myproject/data/regsva_new_z_col.Rdata')
```
## method
* scale the microarrary expression data for each sample to z-score accross all genes
* do ssgsea on the converted microarray expression data
* compute correlation scores for ssgsea score to the drug response data (IC50)

## Result 1: drug correlation analysis with targets sets 
### pdf figure with dendogram
```{r drugdata input}
sample.info <- read.csv('/Users/ll/Desktop/git/myproject/data/CCLE/sample_info.csv')
sample.accession<-sample.info$Accession
drugdata<-read.csv('/Users/ll/Desktop/git/myproject/data/CCLE/CCLE_NP24.2009_Drug_data_2015.02.24.csv')
drugdata.cast<-cast(drugdata,formula = Primary.Cell.Line.Name ~ Compound, value = 'IC50..uM.')
row.names(drugdata.cast)<-drugdata.cast$Primary.Cell.Line.Name
drugdata.cast$Primary.Cell.Line.Name<-NULL
drugdata.cast<-as.data.frame(drugdata.cast)
drug.cl<-row.names(drugdata.cast)
sample.info.cl<-as.character(sample.info$Title)
drug.pos<-match(drug.cl,sample.info.cl)
drug.pos<-drug.pos[!is.na(drug.pos)]
drug.cl.accession<-as.character(sample.accession[drug.pos])
drug.cl.tocorre<-match(drug.cl.accession,colnames(regsva))
drug.cl.tocorre<-drug.cl.tocorre[!is.na(drug.cl.tocorre)]
tocorre.ssgsea<-regsva[,drug.cl.tocorre]
tocorre.col.name<-sample.info$Title[match(colnames(tocorre.ssgsea),sample.accession)]
colnames(tocorre.ssgsea)<-tocorre.col.name
tocorre.drug.pos<-match(colnames(tocorre.ssgsea),drug.cl)
tocorre.drug.pos<-tocorre.drug.pos[!is.na(tocorre.drug.pos)]
tocorre.drug<-drugdata.cast[tocorre.drug.pos,]
```
```{r,warning=FALSE}
regulator.drug.ssgsea.cor<-apply(tocorre.drug,2,function(x){
  ss <- x
  result<-apply(tocorre.ssgsea,1,function(x){return(cor.test(x,ss,method = 'spearman')$estimate)})
  return(result)
    }
  )
```
```{r class}
regulator.class<- read.delim('/Users/ll/Desktop/git/myproject/data/Gene_regulators/Regulator classification.txt',row.names = 1)
druginfo<- read.csv('/Users/ll/Desktop/git/myproject/data/CCLE/CCLE_NP24.2009_profiling_2012.02.20.csv',row.names = 1)
geneage<-read.csv('/Users/ll/Desktop/git/myproject/data/Gene_regulators/geneIDs_entrez_final_phylostrata_phy1_phy2_phy3_no_TCGA.txt',header = TRUE,row.names = 1)
```
```{r gene}
hp.gene<-row.names(regulator.drug.ssgsea.cor)
toremove.gene <- hp.gene[which(regulator.class[hp.gene,1] %in% c('None',NA))]
regulator.drug.ssgsea.cor.sub<-subset(regulator.drug.ssgsea.cor, !hp.gene %in% toremove.gene)
```
```{r graphics,echo = FALSE,out.width = 960,out.height=960}
knitr::include_graphics('drugheatmap_target.pdf')
```
### interactive figure without dendogram (same order)
```{r info,include = FALSE}
my_palette <- colorRampPalette(c("dodgerblue3", "white", "deeppink3"))(n = 4)
#pdf('drugheatmap_target_nodup.pdf',width = 12,height = 12) 
hp<-heatmap.2(regulator.drug.ssgsea.cor.sub,trace = 'none',breaks = c(-1,-0.5,0,0.5,1),col = my_palette)
#dev.off()
new.row.name<-row.names(regulator.drug.ssgsea.cor.sub)[rev(hp$rowInd)]
new.col.name<-colnames(regulator.drug.ssgsea.cor.sub)[hp$colInd]
regulator.drug.ssgsea.cor.2<-regulator.drug.ssgsea.cor.sub[new.row.name,new.col.name]
#heatmap.2(regulator.drug.ssgsea.cor.2,Rowv = FALSE, Colv = FALSE,trace = 'none',breaks = c(-1,-0.5,0,0.5,1),col = my_palette)
```
```{r graphics2,echo = FALSE,out.width = 960,out.height=960}
knitr::include_graphics('drugheatmap_target_nodup.pdf')
```
```{r inter,fig.width=10,fig.height=10}
heatmaply(regulator.drug.ssgsea.cor.2,trace = 'none',breaks = c(-1,-0.5,0,0.5,1),col = my_palette,row_dend_left = TRUE, Rowv = FALSE, Colv = FALSE,plot_method = "plotly")
```
```{r drug2,include=FALSE}
drugname<-row.names(druginfo)
drugname[21]<-'Panobinostat'
drugname[4]<-'PF2341066'
drugname[6]<-'ZD-6474'
row.names(druginfo)<-drugname
a <- druginfo['Irinotecan',3] 
druginfo['Irinotecan',3]<- druginfo['Irinotecan',4]
druginfo['Irinotecan',4] <- a
```


### interactive figure with target set class labels

```{r class new,fig.width=10,fig.height=10}
hp.class<-regulator.class[new.row.name,]
hp.class<-as.character(hp.class)
heatmaply(regulator.drug.ssgsea.cor.2,trace = 'none',breaks = c(-1,-0.5,0,0.5,1),col = my_palette,row_dend_left = TRUE, Rowv = FALSE, Colv = FALSE,plot_method = "plotly",RowSideColors = hp.class)
```

### interactive figure with regulator age labels
```{r age, echo = FALSE,fig.width=10,fig.height=10}
hp.age <- geneage[new.row.name,2]
uc_group<-ifelse(hp.age<4,'uc','mc')
heatmaply(regulator.drug.ssgsea.cor.2,trace = 'none',breaks = c(-1,-0.5,0,0.5,1),col = my_palette,RowSideColors = uc_group,Rowv = FALSE, Colv = FALSE)
```

## Result 2: the information on the cluster drugs
```{r dendro,echo=FALSE,results='asis'}
hp.dendrogram.col<-hp$colDendrogram
hc <- as.hclust(hp.dendrogram.col)
hc_cut<-cutree( hc, k=5)
hc_cut.order<-hc_cut[order(hc_cut)]
druginfo.2<-druginfo[names(hc_cut.order),]
druginfo.2$group<-hc_cut.order
druginfo.total<-druginfo.2
kable(druginfo.total[,c(2,3,4,7)])
```

## Result 3: drug response correlate with goslim
```{r goslim1,eval=FALSE}
gene.expression.nohemo.z<-read.csv('/Users/ll/Desktop/git/myproject/data/CCLE/geneexpression_zscore.csv',row.names = 1)
mydata<-getGEO(filename='/Users/ll/Desktop/git/myproject/data/GSE36133_series_matrix.txt')
slim_terms <- getOBOCollection("/Users/ll/Desktop/git/myproject/data/Gene_regulators/goslim_generic.obo")
xx <- as.list(org.Hs.egGO2ALLEGS)
goids <- xx[2:3]
gonames<-Term(GOTERM)
go_bp_children <- GOBPOFFSPRING$"GO:0008150"
slim_terms <- ids(slim_terms)[which(ids(slim_terms) %in% go_bp_children)]
slim_terms <- cbind(slim_terms, unname(gonames[match(slim_terms, names(gonames))]))
pathways <- xx[slim_terms[,1]]
pathways <- pathways[!is.na(names(pathways))]
names(pathways) <- unname(gonames[match(names(pathways), names(gonames))])
anno <- fData(mydata)
ent<-as.character(anno$ORF)
genename<-getSYMBOL(ent,data='org.Hs.eg')
anno <- cbind(anno,genename)
gene.expression.nohemo.z<-gene.expression.nohemo.z[complete.cases(anno),]
anno<-anno[complete.cases(anno),]
rownames(gene.expression.nohemo.z)<-anno$ORF
gogsva<-gsva(data.matrix(gene.expression.nohemo.z),pathways,method=c('ssgsea'))
save(gogsva,file ='../data/gogsva_z_col.Rdata')
```

```{r goslim2}
load(file ='/Users/ll/Desktop/git/myproject/data/gogsva_z_col.Rdata')
tocorre.goslim<-gogsva[,drug.cl.tocorre]
```
```{r goslim3,warning=FALSE}
regulator.drug.goslim.cor<-apply(tocorre.drug,2,function(x){
  ss <- x
  result<-apply(tocorre.goslim,1,function(x){return(cor.test(x,ss,method = 'spearman')$estimate)})
  return(result)
    }
  )
```
```{r goslim4,include = FALSE,fig.width=12,fig.height=12}
my_palette <- colorRampPalette(c("dodgerblue3", "white", "deeppink3"))(n = 4)
#pdf('drugheatmap_goslim.pdf',width = 12,height = 12) 
hp.2<-heatmap.2(regulator.drug.goslim.cor,trace = 'none',breaks = c(-1,-0.5,0,0.5,1),col = my_palette)
#dev.off()
```

```{r goslim5}
new.row.name<-row.names(regulator.drug.goslim.cor)[rev(hp.2$rowInd)]
new.col.name<-colnames(regulator.drug.goslim.cor)[hp.2$colInd]
regulator.drug.goslim.cor.2<-regulator.drug.goslim.cor[new.row.name,new.col.name]
```

```{r graphics 2,echo = FALSE,out.width = 960,out.height=960}
knitr::include_graphics('drugheatmap_goslim.pdf')
```

###interactive plot
```{r goslim6,fig.width=10,fig.height=10}
heatmaply(regulator.drug.goslim.cor.2,trace = 'none',breaks = c(-1,-0.5,0,0.5,1),col = my_palette,Rowv = FALSE, Colv = FALSE)
```
```{r goslim7}
hp.2.dendrogram.col<-hp.2$colDendrogram
hc.2 <- as.hclust(hp.2.dendrogram.col)
hc_cut.2<-cutree( hc.2, k=5)
hc_cut.2.order<-hc_cut.2[order(hc_cut.2)]
druginfo.3<-druginfo[names(hc_cut.2.order),]
druginfo.3$group<-hc_cut.2.order
druginfo.total.2<-druginfo.3
kable(druginfo.total.2[,c(2,3,4,7)])
```












