---
title: "ccle drug correlation"
output:
  pdf_document: default
  html_notebook: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE)
library(reshape)
library(GEOquery)
library(gplots)
library(dplyr)
library(GSVA)
library(GSEABase)
library(knitr)
library(org.Hs.eg.db)
library(annotate)
library(data.table)
library(biomaRt)
library(GO.db)
```

```{r load data}
load('./data/selected_cl.Rdata')
load('./data/regulators_5_cl_20_tar.Rdata')
load('./data/regulator_regulation.Rdata')
```

```{r 0}
ccle.expression<-fread('./data/CCLE/CCLE_Expression_Entrez_2012-09-29.gct',skip = 2, sep = '\t')
ccle.expression_o <- ccle.expression
ccle.expression <- as.data.frame(ccle.expression)
ccle.expression <- ccle.expression[,-c(1,2)]
ccle.anno <- ccle.expression_o[,c(1,2)]
toremove <- which(ccle.anno$Description == '')
ccle.anno <- ccle.anno[-toremove,]
ccle.expression<-ccle.expression[-toremove,]
#gene.exp.z <- apply(ccle.expression,2,scale)
#colnames(gene.exp.z) <- colnames(ccle.expression)
#save(gene.exp.z,file = '../data/CCLE/geneexpression_zscore_new.Rdata' )
#write.csv(gene.exp.z,file = '../data/CCLE/geneexpression_zscore_new.csv')
```

```{r sample}
sample.info<-fread('./data/CCLE/CCLE_sample_info_file_2012-10-18.txt', sep = '\t')
sample.info <- as.data.frame(sample.info)
cl.pos<-match(colnames(ccle.expression),sample.info$`CCLE name`)
cl.info <- sample.info[cl.pos,]
cl.toremove <- which(cl.info$`Site Primary` == 'haematopoietic_and_lymphoid_tissue')
ccle.exp <- ccle.expression[,cl.tokeep]
```


```{r input}
gene.expression.z<-read.csv('./data/CCLE/geneexpression_zscore_new.csv',row.names = 1)
load('./data/ccle_regsva_regulator.Rdata')
```

```{r drug}
drugdata<-read.csv('./data/CCLE/CCLE_NP24.2009_Drug_data_2015.02.24.csv')
drugdata.cast<-cast(drugdata,formula = CCLE.Cell.Line.Name ~ Compound, value = 'IC50..uM.')
row.names(drugdata.cast)<-drugdata.cast$CCLE.Cell.Line.Name
drugdata.cast$CCLE.Cell.Line.Name<-NULL
drugdata.cast<-as.data.frame(drugdata.cast)
```

```{r drug 2}

cl.tocorre <- intersect(colnames(ccle.exp),row.names(drugdata.cast))

position <- match(selected.regulator,ccle.anno$Description)
toremove<- is.na(position)
regulator.tokeep <- selected.regulator[!toremove]
position <- match(regulator.tokeep,ccle.anno$Description)
cl.exp <- ccle.exp[position,]
row.names(cl.exp) <- regulator.tokeep
exp.tocorre <- cl.exp
ssgsea.tocorre <- regsva[selected.regulator,cl.tocorre]
tocorre.drug <- drugdata.cast[cl.tocorre,]
```



```{r corr1}
regulator.drug.ssgsea.cor<-apply(tocorre.drug,2,function(x){
  names(x) <- cl.tocorre
  cl.sen <- cl.tocorre[x < 8]
  cl.sen <- cl.sen[! is.na(cl.sen)]
  ss <- x[cl.sen]
  result<-apply(ssgsea.tocorre[,cl.sen],1,function(x){return(cor.test(x,ss,method = 'spearman')$estimate)})
  return(result)
    }
  ) # how much is the number of cl per correlation experiment
```

```{r corr2}

regulator.drug.exp.cor<-apply(tocorre.drug,2,function(x){
  names(x) <- cl.tocorre
  cl.sen <- cl.tocorre[x < 8]
  cl.sen <- cl.sen[! is.na(cl.sen)]
  ss <- x[cl.sen]
  result<-apply(exp.tocorre[,cl.sen],1,function(x){return(cor.test(x,ss,method = 'spearman')$estimate)})
  return(result)
    }
  ) # how much is the number of cl per correlation experiment
```


## drug correlation analysis with regulator expression   
* spearman correlation of the IC50 values of each drug to the expression of regulators   
* negative correlation means sensitive, positive correlation means resistant   
* only response cell lines(IC50<8)  are used in the analysis   

```{r plot1,fig.height=10,fig.width=10}
#pdf('cancer_drug_cor_exp.pdf',width = 12,height = 12)
my_palette <- colorRampPalette(c("dodgerblue3", "white", "deeppink3"))(n = 4)
heatmap.2(regulator.drug.exp.cor,main = 'Correlation coefficients analysis of \n cancer drugs for regulator expression',
          dendrogram="both",
          breaks = c(-1,-0.5,0,0.5,1),
          col = my_palette,trace="none",
          keysize = 0.7,
          #seriate = 'mean',
          #row_dend_left = TRUE,
          #plot_method = 'plotly'
          margins = c(10,20),
          cexRow = 0.8,
          density.info = 'none',
          cexCol = 1
          #Colv = FALSE,
          )

```

## Number of response cell lines of each drug

```{r corr3}
regulator.drug.ssgsea.num<-apply(tocorre.drug,2,function(x){
  names(x) <- cl.tocorre
  cl.sen <- cl.tocorre[x < 8]
  cl.sen <- cl.sen[! is.na(cl.sen)]
  ss <- x[cl.sen]
  return(length(ss))
    }
  ) 
barplot(regulator.drug.ssgsea.num,names.arg = colnames(tocorre.drug),las = 2,ylab = 'number of response cell lines')
```

## drug correlation analysis with target gene set expression   
* spearman correlation of the IC50 values of each drug to the expression of target gene sets (single sample GSEA score)   
* negative correlation means sensitive, positive correlation means resistant   
* only response cell lines(IC50<8)  are used in the analysis   

```{r heatmap2,fig.height=10,fig.width=10}

heatmap.2(regulator.drug.ssgsea.cor,main = 'Correlation coefficients analysis of \n cancer drugs for target set expression',
          dendrogram="both",
          breaks = c(-1,-0.5,0,0.5,1),
          col = my_palette,trace="none",
          keysize = 0.7,
          #seriate = 'mean',
          #row_dend_left = TRUE,
          #plot_method = 'plotly'
          margins = c(10,20),
          cexRow = 0.8,
          density.info = 'none',
          cexCol = 1
          #Colv = FALSE,
          )


```


```{r reg class}
regulator.class<- read.delim('./data/Gene_regulators/Regulator classification.txt',row.names = 1)
druginfo<- read.csv('./data/CCLE/CCLE_NP24.2009_profiling_2012.02.20.csv',row.names = 1)
geneage<-read.csv('./data/Gene_regulators/geneIDs_entrez_final_phylostrata_phy1_phy2_phy3_no_TCGA.txt',header = TRUE,row.names = 1)
```

## drug correlation analysis with GO slim biological processes gene set expression   
* spearman correlation of the IC50 values of each drug to the expression of GO slim biological processes gene sets (single sample GSEA score)   
* negative correlation means sensitive, positive correlation means resistant   
* only response cell lines(IC50<8)  are used in the analysis   


```{r goslim2}
load(file ='./data/gosva.Rdata')
tocorre.goslim<-gogsva[,cl.tocorre]
```

```{r corr4}
regulator.drug.goslim.cor<-apply(tocorre.drug,2,function(x){
  ss <- x
  names(x) <- cl.tocorre
  cl.sen <- cl.tocorre[x < 8]
  cl.sen <- cl.sen[! is.na(cl.sen)]
  ss <- x[cl.sen]
  result<-apply(tocorre.goslim[,cl.sen],1,function(x){return(cor.test(x,ss,method = 'spearman')$estimate)})
  return(result)
    }
  )
```

```{r heatmap3,fig.height=10,fig.width=10}
load("./data/goslim_order.Rdata")
heatmap.2(regulator.drug.goslim.cor[order.2,],
          main = 'Correlation coefficients analysis of \n cancer drugs for cellular \n processes expression',
          Rowv = FALSE,
          dendrogram="column",
          breaks = c(-1,-0.5,0,0.5,1),
          col = my_palette,trace="none",
          keysize = 0.7,
          #seriate = 'mean',
          #row_dend_left = TRUE,
          #plot_method = 'plotly'
          margins = c(10,20),
          cexRow = 0.8,
          density.info = 'none',
          cexCol = 1
          #Colv = FALSE,
          )
```

### same order as previous analysis

```{r heatmap4,fig.height=10,fig.width=10}
load("./data/goslim_order.Rdata")
my_palette <- colorRampPalette(c("dodgerblue3", "white", "deeppink3"))(n = 4)
hp.2<-heatmap.2(regulator.drug.goslim.cor[order.2,],trace = 'none',breaks = c(-1,-0.5,0,0.5,1),col = my_palette)
regulator.key <- selected.regulator[regulator_tb$class == 'CNV,target set']
heatmap.2(regulator.drug.goslim.cor[order.2,],Rowv = FALSE,trace = 'none',breaks = c(-1,-0.5,0,0.5,1),col = my_palette,density.info = 'none')

```

## Cluster of drugs from GO slim correlation analysis (shown in column 'group')

```{r goslim7}
hp.2.dendrogram.col<-hp.2$colDendrogram
hc.2 <- as.hclust(hp.2.dendrogram.col)
hc_cut.2<-cutree( hc.2, k=2)
hc_cut.2.order<-hc_cut.2[order(hc_cut.2)]
names(hc_cut.2.order)[which(names(hc_cut.2.order) == 'ZD-6474')]<-'Vandetanib'
druginfo.3<-druginfo[names(hc_cut.2.order),]
druginfo.3$group<-hc_cut.2.order
druginfo.total.2<-druginfo.3
kable(druginfo.total.2[,c(2,3,4,7)])
```
















