---
title: "Drug response"
output:
  html_document: default
  html_notebook: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(reshape)
library(GEOquery)
library(gplots)
library("dplyr")
```

```{r regulator,include = FALSE}
mydata<-getGEO(filename='/Users/ll/Desktop/work/data/GSE36133_series_matrix.txt')
mutation_o<-read.delim('/Users/ll/Desktop/work/data/CCLE/CCLE_DepMap_18Q1_maf_20180207 (1).txt')
mutation<-mutation_o
mutation_s<-mutation_o[which(is.na(mutation_o$ExAC_AF)),]
mutation_s<-mutation_s[which(mutation_s$isDeleterious == 'TRUE'),]
mutation_s_w<-mutation_s
mutation_s_w<-data.frame(mutation_s$Hugo_Symbol,mutation_s$Tumor_Sample_Barcode)
genelist<-as.character(unique(mutation_s_w[,1]))
rs<-function(x){
  c<-as.character(mutation_s_w[which(mutation_s_w$mutation_s.Hugo_Symbol==x),2])
  return(c)
  }
a<-lapply(genelist,rs)
names(a)<-genelist
mutatedcelllines<-a
b<-lapply(mutatedcelllines,length)
mutatedcelllines_num<-as.data.frame(b)
mutatedcelllines_num<-t(mutatedcelllines_num)
colnames(mutatedcelllines_num)<-c('freq')

regulator<-read.delim('/Users/ll/Desktop/work/analysis/analysis_on_mutation/Gene_regulators/regulator classification.txt')
regulator<-regulator[,1]
regulator<-as.character(regulator)
regulator<-unique(regulator)

reg <- read.delim('/Users/ll/Desktop/work/analysis/analysis_on_mutation/Gene_regulators/Regulatory network.txt')
regulator <- unique(reg[,1])
target <- unique(reg[,2])
reg_m<-matrix(nrow=length(regulator),ncol=length(target))
colnames(reg_m)<-target
row.names(reg_m)<-regulator
reg_m_n<-reg_m
for (i in 1:length(reg[,1])){
  
  reg_m_n[reg[i,1],reg[i,2]] = 1
}

#count targets for each regulator
reg_m_n[is.na(reg_m_n)]<-0
reg_tar<-apply(reg_m_n,1,sum)

info<-read.delim('/Users/ll/Desktop/work/data/CCLE/CCLE_sample_info_file_2012-10-18.txt')
celllines<-unique(info$CCLE.name)
celllines<-as.character(celllines)
```
```{r info,include=FALSE}
sample.info <- read.csv('/Users/ll/Desktop/work/data/CCLE/sample_info.csv')
cl.sample<-c()
n<-1
k <- 1
t <- c()
sample.title<-as.character(sample.info$Title)
for (i in sample.title){
a<- as.character(info$CCLE.name[info$Cell.line.primary.name == i])
if(length(a) == 0){
  cl.sample[n]<-i
} 
else{
cl.sample[n]<- a}
  n <-n + 1
  }
sample.accession<-sample.info$Accession
sample.info.correct<-rbind(as.character(sample.accession),cl.sample)
sample.info.correct<-t(sample.info.correct)

cl<-rep(0,length(cl.sample))
regulator<-as.character(regulator)
regulators.cl<-regulator[which(regulator %in% names(mutatedcelllines))]
n <- length(cl.sample)
cl[n+1] <- 'cl'
for (i in regulators.cl){
cl.matrix <- cl.sample %in% mutatedcelllines[[i]]
cl.matrix[n+1] <- i 
cl<-rbind(cl.matrix,cl)
}
row.names(cl)<-as.vector(cl[,ncol(cl)])
cl.o<-cl
cl<-cl[,-ncol(cl)]
colnames(cl)<-cl.sample
cl<-cl[-which(row.names(cl) == 'cl'),]
gene.nomutate<-c()
i<-1
pd<-pData(mydata)
toremove<-as.character(pd$`primary site:ch1`)
toremove<-toremove == 'haematopoietic_and_lymphoid_tissue'
cl_nohemo<-cl[,-which(toremove == TRUE)]
for (gene in regulators.cl){
  if (sum(as.numeric(as.logical(cl[gene,]))) == 0){
    gene.nomutate[i]<-gene
    i<-i+1
  }
}
cl.m <- apply(cl,1,as.logical)
reg.10 <- names(reg_tar[reg_tar>10])
regulators.cl.mutate<-regulators.cl[-which(regulators.cl %in% gene.nomutate)]
reg_cl <- apply(cl.m,2,sum)
reg_5 <- names(reg_cl[reg_cl>3])
genes.toanalyze<-intersect(reg.10,reg_5)
genes.toanalyze<-intersect(genes.toanalyze,regulators.cl.mutate)
```

```{r input, include = FALSE}
drugdata<-read.csv('~/Desktop/work/data/CCLE/CCLE_NP24.2009_Drug_data_2015.02.24.csv')
drugdata.cast<-cast(drugdata,formula = CCLE.Cell.Line.Name ~ Compound, value = 'IC50..uM.')
row.names(drugdata.cast)<-drugdata.cast$CCLE.Cell.Line.Name
drugdata.cast$CCLE.Cell.Line.Name<-NULL
drugdata.cast<-as.data.frame(drugdata.cast)
drugdata.cast.t<-t(drugdata.cast)
drug.cl<-colnames(drugdata.cast.t)
regulator.cl<-cl[,which(colnames(cl)%in%drug.cl)]
hemocl<-colnames(cl)[toremove]
regulator.cl.nohemo<-regulator.cl[,-which(colnames(regulator.cl)%in%hemocl)]
regulator.cl.nohemo.filter<-regulator.cl.nohemo[genes.toanalyze,]
```

## Create Cell line sets

```{r cell line set, include= FALSE}
# create gmt for ssgsea
#cl.list<-list()
#cl.name<-colnames(regulator.cl.nohemo.filter)
#for (i in (1:nrow(regulator.cl.nohemo.filter))){
#  cl.list[[i]]<-cl.name[which(as.logical(regulator.cl.nohemo.filter[i,]))]
#}
#names(cl.list)<-rownames(regulator.cl.nohemo.filter)
#lapply(cl.list, function(x) write.table(t(as.data.frame(x)), 'test.tsv'  , append= T, sep='\t' ))
#library(stringi)
#res <- as.data.frame(t(stri_list2matrix(cl.list)))
#colnames(res) <- unique(unlist(sapply(cl.list, names)))
#row.names(res)<-rownames(regulator.cl.nohemo.filter)
#write.table(res,file = 'clset.gmt',sep = '\t')
```

```{r ssgsea, include = FALSE}
library(GSVA)
library(GSEABase)
clset <- getGmt('/Users/ll/Desktop/work/analysis/analysis_on_expression/gset/clset.gmt')
####super slow!!!
#regsva<-gsva(e[,-which(toremove==TRUE)],regset,method=c('ssgsea'))
```

```{r wilcox, include=FALSE}
cl.number<-apply(regulator.cl.nohemo.filter,1,function(x){return(sum(as.logical(x)))})
tokeep<- cl.number > 0
regulator.cl.nohemo.filter.2<-regulator.cl.nohemo.filter[tokeep,]
cl.name<-colnames(regulator.cl.nohemo.filter.2)
wilcox.drug<-function(x){
  cl.logic<-regulator.cl.nohemo.filter.2[x,]
  cl.mutate<-cl.name[as.logical(cl.logic)]
  cl.notmutate<-cl.name[-as.logical(cl.logic)]
  a<-sapply(c(1:nrow(drugdata.cast.t)),
           function(x){
               cl.mutate.drug<-drugdata.cast.t[x,cl.mutate]
               cl.notmutate.drug<-drugdata.cast.t[x,cl.notmutate]
               if(length(cl.mutate.drug[!is.na(cl.mutate.drug)])>=1){
                  r<-wilcox.test(cl.mutate.drug,cl.notmutate.drug,alternative = 'less')
                  return(r$p.value)}
               else{
                  return(NA)
               }
           }
               )
  names(a)<-row.names(drugdata.cast.t)
  a.adjust<-p.adjust(a,method = 'BH')
  return(a.adjust)
  #return(a)
}
drug.cl.efficacy.wilcox <- data.frame(matrix(nrow = nrow(drugdata.cast.t),ncol = nrow(regulator.cl.nohemo.filter.2)))
row.names(drug.cl.efficacy.wilcox)<-row.names(drugdata.cast.t)
colnames(drug.cl.efficacy.wilcox)<- row.names(regulator.cl.nohemo.filter.2)
for (i in c(1:nrow(regulator.cl.nohemo.filter.2))){
  drug.cl.efficacy.wilcox[,i]<-wilcox.drug(i)}
```

## Heatmap for Wilcox rank sum test p-value (red for siginificant test, white is NA)
```{r heatmap,fig.width=40,fig.height=20, echo = FALSE}
toheatmap<-data.matrix(drug.cl.efficacy.wilcox)
my_palette <- colorRampPalette(c("blue", "white"))(n = 2)
heatmap.2(toheatmap,main = 'drug efficacywlicox  test',dendrogram="both",breaks = c(0,0.05,1),trace="none",density.info = 'density',key.title = 'P-value for \ndrug efficacy wlicox  test result',keysize = 0.6,margins = c(3,20),cexRow = 1.5,cexCol = 1.2)
```
```{r analysis}
#drug.cl.efficacy.wilcox[with(drug.cl.efficacy.wilcox,drug.cl.efficacy.wilcox<0.05)]
sig.drug.pos<-which(as.matrix(drug.cl.efficacy.wilcox)<0.05, arr.ind=TRUE)
sig.drug.df<-as.data.frame(sig.drug.pos)
sig.drug.df$p.value<-drug.cl.efficacy.wilcox[which(as.matrix(drug.cl.efficacy.wilcox)<0.05, arr.ind=TRUE)]
sig.drug.df$row<-row.names(drug.cl.efficacy.wilcox)[sig.drug.df$row]
sig.drug.df$col<-colnames(drug.cl.efficacy.wilcox)[sig.drug.df$col]
colnames(sig.drug.df)<-c('drug','regulator','pvalue.adjusted')
drug<-as.character(unique(drugdata$Compound))
drug_target<-drugdata[which(drugdata$Compound %in% drug),c('Compound','Target')]
target<-as.character(unique(drugdata$Target))
drug_target<-distinct(drug_target)
sig.drug.df$target <- drug_target$Target[match(sig.drug.df$drug,drug_target$Compound)]
sig.drug.df
#sig.drug<-list()
#drug.index<-unique(sig.drug.pos[,1])
#for(i in c(1:length(drug.index))){
#  sig.drug.col<-sig.drug.pos[which(sig.drug.pos[,1] == drug.index[i]),2]
#  sig.drug[[i]] <-drug.cl.efficacy.wilcox[drug.index[i],sig.drug.col]
#  if(length(sig.drug[[i]]) == 1){names(sig.drug[[i]])<-colnames(drug.cl.efficacy.wilcox[sig.drug.col])}
#  }
#names(sig.drug)<-row.names(drug.cl.efficacy.wilcox)[drug.index]
```

## Drug information
Information:
    
    * AZD6244(Selumetinib), treatmentfor NSCLC, thyroid cancer. Mechanism: block subtype MEK1, MEK2 of the MAPK kinase, downstream of BRAF.
    
    * PD-0325901 MEK inhibitor
    
    * L-685458  Inhibitor for gamma-Secretase
    
    * PLX4720  Raf Inhibitor
    
    * LBW242 Inhibitor of Apoptosis Protein (IAP) inhibitor 
    
    * Nutlin-3 Inhibitor of MDM2
    
    * Saracatinib(AZD0530) Inhibitor of the Src family of tyrosine kinases
    
    * Palbociclib(PD-0332991) Inhibitor of CDK4/6
    
    * PHA665752  c-Met inhibitor C-Met/HGF/SF signaling
    
    * Sorafenib small inhibitor of several tyrosine protein kinases, such as VEGFR, PDGFR and Raf family kinases 
    
    * Dovitinib (TKI-258) multitargeted RTK inhibitor
    
    * Crizotinib (PF2341066) a novel inhibitor of anaplastic lymphoma kinase and c-Met
    approved for treatment of some non-small cell lung carcinoma (NSCLC) in the US and some other countries, and undergoing clinical trials testing its safety and efficacy in anaplastic large cell lymphoma, neuroblastoma, and other advanced solid tumors in both adults and children.
    
    * Topotecan chemotherapy drug which can be used to treat ovarian cancer, small cell lung cancer and cervical cancer, topoisomerase-I inhibitor
    
```{r correlation heatmap}
##
```

