---
title: "gene set analysis"
author: "bingrus"
date: "2018-08-14"
output: workflowr::wflow_html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(htmlwidgets)
library(limma)
library(org.Hs.eg.db)
library(annotate)
library(GEOquery)
library(grid)
library(ggplot2)
library(reshape)
library(GSVA)
library(GSEABase)
library(GO.db)
library(gplots)
library(heatmaply)
```

```{r regulator,include = FALSE}
mydata<-getGEO(filename='/Users/ll/Desktop/work/data/GSE36133_series_matrix.txt')
mutation_o<-read.delim('/Users/ll/Desktop/work/data/CCLE/CCLE_DepMap_18Q1_maf_20180207 (1).txt')
mutation<-mutation_o
mutation_s<-mutation_o[which(is.na(mutation_o$ExAC_AF)),]
mutation_s<-mutation_s[which(mutation_s$isDeleterious == 'TRUE'),]
mutation_s_w<-mutation_s
mutation_s_w<-data.frame(mutation_s$Hugo_Symbol,mutation_s$Tumor_Sample_Barcode)
genelist<-as.character(unique(mutation_s_w[,1]))
rs<-function(x){
  c<-as.character(mutation_s_w[which(mutation_s_w$mutation_s.Hugo_Symbol==x),2])
  return(c)
  }
a<-lapply(genelist,rs)
names(a)<-genelist
mutatedcelllines<-a
b<-lapply(mutatedcelllines,length)
mutatedcelllines_num<-as.data.frame(b)
mutatedcelllines_num<-t(mutatedcelllines_num)
colnames(mutatedcelllines_num)<-c('freq')

regulator<-read.delim('/Users/ll/Desktop/work/analysis/analysis_on_mutation/Gene_regulators/regulator classification.txt')
regulator<-regulator[,1]
regulator<-as.character(regulator)
regulator<-unique(regulator)

reg <- read.delim('/Users/ll/Desktop/work/analysis/analysis_on_mutation/Gene_regulators/Regulatory network.txt')
regulator <- unique(reg[,1])
target <- unique(reg[,2])
reg_m<-matrix(nrow=length(regulator),ncol=length(target))
colnames(reg_m)<-target
row.names(reg_m)<-regulator
reg_m_n<-reg_m
for (i in 1:length(reg[,1])){
  
  reg_m_n[reg[i,1],reg[i,2]] = 1
}

#count targets for each regulator
reg_m_n[is.na(reg_m_n)]<-0
reg_tar<-apply(reg_m_n,1,sum)

info<-read.delim('/Users/ll/Desktop/work/data/CCLE/CCLE_sample_info_file_2012-10-18.txt')
celllines<-unique(info$CCLE.name)
celllines<-as.character(celllines)
```

```{r info,include=FALSE}
sample.info <- read.csv('/Users/ll/Desktop/work/data/CCLE/sample_info.csv')
cl.sample<-c()
n<-1
k <- 1
t <- c()
sample.title<-as.character(sample.info$Title)
for (i in sample.title){
a<- as.character(info$CCLE.name[info$Cell.line.primary.name == i])
if(length(a) == 0){
  cl.sample[n]<-i
} 
else{
cl.sample[n]<- a}
  n <-n + 1
  }
sample.accession<-sample.info$Accession
sample.info.correct<-rbind(as.character(sample.accession),cl.sample)
sample.info.correct<-t(sample.info.correct)

cl<-rep(0,length(cl.sample))
regulator<-as.character(regulator)
regulators.cl<-regulator[which(regulator %in% names(mutatedcelllines))]
n <- length(cl.sample)
cl[n+1] <- 'cl'
for (i in regulators.cl){
cl.matrix <- cl.sample %in% mutatedcelllines[[i]]
cl.matrix[n+1] <- i 
cl<-rbind(cl.matrix,cl)
}
row.names(cl)<-as.vector(cl[,ncol(cl)])
cl.o<-cl
cl<-cl[,-ncol(cl)]
colnames(cl)<-cl.sample
cl<-cl[-which(row.names(cl) == 'cl'),]
gene.nomutate<-c()
i<-1
pd<-pData(mydata)
toremove<-as.character(pd$`primary site:ch1`)
toremove<-toremove == 'haematopoietic_and_lymphoid_tissue'
cl<-cl[,-which(toremove == TRUE)]
for (gene in regulators.cl){
  if (sum(as.numeric(as.logical(cl[gene,]))) == 0){
    gene.nomutate[i]<-gene
    i<-i+1
  }
}
cl.m <- apply(cl,1,as.logical)
reg.10 <- names(reg_tar[reg_tar>10])
regulators.cl.mutate<-regulators.cl[-which(regulators.cl %in% gene.nomutate)]
reg_cl <- apply(cl.m,2,sum)
reg_5 <- names(reg_cl[reg_cl>3])
genes.toanalyze<-intersect(reg.10,reg_5)
genes.toanalyze<-intersect(genes.toanalyze,regulators.cl.mutate)
```

##Camera analysis on biological processes
* select regulators for mutated in at least 3 cell lines with at least 10 targets
* 173 regulators are selected
* Camera gene set analysis conducted on the pairs of mutated cell lines versus non-mutated cell lines for the selected regulators
* Plot the significance level for each competitive gene set test 
    - Plot 1 - P-value 
    - Positive for Up Direction; Negative for Down Direction)
```{r getannotation,include=FALSE}
anno <- fData(mydata)
library(org.Hs.eg.db)
library(annotate)
ent<-as.character(anno$ORF)
genename<-getSYMBOL(ent,data='org.Hs.eg')
anno <- cbind(anno,genename)
gene.ORF<-anno$ORF
gene.name<-anno$genename
```

```{r idx, include = FALSE}
slim_terms <- getOBOCollection("~/Desktop/work/analysis/analysis_on_expression/goslim_generic.obo")
xx <- as.list(org.Hs.egGO2ALLEGS)
goids <- xx[2:3]
gonames<-Term(GOTERM)
go_bp_children <- GOBPOFFSPRING$"GO:0008150"
slim_terms <- ids(slim_terms)[which(ids(slim_terms) %in% go_bp_children)]
slim_terms <- cbind(slim_terms, unname(gonames[match(slim_terms, names(gonames))]))
pathways <- xx[slim_terms[,1]]
pathways <- pathways[!is.na(names(pathways))]
names(pathways) <- unname(gonames[match(names(pathways), names(gonames))])
#gogsva<-gsva(e,pathways,method=c('ssgsea'))
ids<-ids2indices(pathways,gene.ORF)
```

```{r camera , include = FALSE}
camera.targetset<-function(x){
gene<-x
n<-ncol(cl)
cl.design <- matrix(nrow=n,ncol=2)
cl.design[,1] <- as.numeric(as.logical(cl[gene,]))
cl.design.2<-cl.design[,1]
cl.design.2[which(cl.design[,1] == 1)]<- 0
cl.design.2[which(cl.design[,1] == 0)]<- 1
cl.design[,2]<-cl.design.2
colnames(cl.design)<-c('mutated','non_mutated')
r1 <- camera(exprs(mydata)[,-which(toremove == TRUE)], ids, cl.design,contrast=makeContrasts(mutated - non_mutated, levels=cl.design))
#Pvalue<-r1$Pvalue
#Direction<-r1$Direction 
#Pvalue<-1-Pvalue
#Pvalue[which(Direction == 'Down')]<- -Pvalue[which(Direction == 'Down')]
return(r1)}
```

```{r camera2,include=FALSE}
#result.camera<-lapply(genes.toanalyze,camera.targetset)
#names(result.camera)<-genes.toanalyze
#save(result.camera,file = '/Users/ll/Desktop/work/analysis/analysis_on_expression/goslim_camera.Rdata')
```

## Go Slim no reorder (blue for mc, red for uc)
```{r camera to table,echo=FALSE,fig.height=80,fig.width=120}
load('/Users/ll/Desktop/work/analysis/analysis_on_expression/goslim_camera.Rdata')
process.name<-names(pathways)
geneage<-read.csv('/Users/ll/Desktop/work/analysis/analysis_on_mutation/Gene_regulators/geneIDs_entrez_final_phylostrata_phy1_phy2_phy3_no_TCGA.txt',header = TRUE,row.names = 1)
mode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}
convertgeneage<-function(x){
  process.gene<-pathways[[x]]
  process.age<-geneage[with(geneage,geneage$Entrez %in% process.gene),2]
  return(mode(process.age))
}
processage<-sapply(c(1:69),convertgeneage)
names(processage)<-process.name
uc_group<-ifelse(processage<4,'uc','mc')
uc_color<-ifelse(processage<4,'pink','blue')
process.name.sort<-names(sort(uc_color))
tomatch<-process.name.sort
converttopvalue<-function(x){
#  a<-result.camera[[x]]
a<-result.camera[[x]]
Pvalue<-1-a$PValue
Direction<-a$Direction
Pvalue[ifelse(Direction == 'Down',TRUE,FALSE)]<--Pvalue[ifelse(Direction == 'Down',TRUE,FALSE)]
position <- match(tomatch,row.names(a))
Pvalue.new<-Pvalue[position]
names(Pvalue.new)<-tomatch
return(Pvalue.new)
}
camera.matrix<-sapply(c(1:length(result.camera)),converttopvalue)
colnames(camera.matrix)<-names(result.camera)
toheatmap<-data.matrix(camera.matrix)
my_palette <- colorRampPalette(c("blue", "white", "red"))(n = 4)
heatmap.2(toheatmap,main = 'Camera gene test result (no reorder)',Rowv = FALSE,dendrogram="col",breaks = c(-1,-0.95,0,0.95,1),col = my_palette,trace="none",RowSideColors = sort(uc_color),density.info = 'density',key.title = '1 - P-value for \ncamera gene set test result',keysize = 0.6,margins = c(3,20),cexRow = 1.5,cexCol = 1.2)
#tiff(file = '/Users/ll/Desktop/work/analysis/analysis_on_expression/figures/Camera gene test result (no reorder).tiff')

#heatmap.2(mat_data,
#  cellnote = mat_data,  # same data set for cell labels
#  main = "Correlation", # heat map title
#  notecol="black",      # change font color of cell labels to black
#  density.info="none",  # turns off density plot inside color legend
#  trace="none",         # turns off trace lines inside the heat map
#  margins =c(12,9),     # widens margins around plot
#  col=my_palette,       # use on color palette defined earlier
#  breaks=col_breaks,    # enable color transition at specified limits
#  dendrogram="row",     # only draw a row dendrogram
#  Colv="NA")   
#converttopvalue(1)
```

```{r interactive 0,echo = FALSE,dpi=300,include = FALSE}
heatmaply(toheatmap,main = 'Camera gene test result (no reorder)',Rowv = FALSE,dendrogram="col",breaks = c(-1,-0.95,0,0.95,1),col = my_palette,trace="none",RowSideColors = sort(uc_color),density.info = 'density',key.title = '1 - P-value for \ncamera gene set test result',keysize = 0.6,cexRow = 0.7,cexCol = 0.3,colorbar_xanchor = 'left',colorbar_yanchor = 'bottom',margins = c(2,210,30,2))
```

```{r age,include = FALSE}
heatmap.order<-heatmap.2(toheatmap,main = 'Camera gene test result (ordered)',Rowv = TRUE,breaks = c(-1,-0.99,-0.95,0,0.95,0.99,1),col="heat.colors",trace="none")
```

## Go Slim reordered (blue for mc, red for uc)
```{r plot,echo = FALSE,fig.width= 120, fig.height= 80}
neworder<-heatmap.order$rowInd
process.name.2<-process.name.sort[rev(neworder)]
#process.name.2
uc_color.2<-uc_color[process.name.2]
uc_group.2<-uc_group[process.name.2]
toheatmap.2<-toheatmap[process.name.2,]
my_palette <- colorRampPalette(c("dodgerblue3", "white", "deeppink3"))(n = 4)
heatmap.2(toheatmap.2,main = 'Camera gene test result (ordered) for Goslim set',Rowv = FALSE,dendrogram = 'col',breaks = c(-1,-0.95,0,0.95,1),col = my_palette,trace="none",RowSideColors = uc_color.2,density.info = 'density',key.title = '1 - P-value for \ncamera gene set test result',keysize = 0.6,margins = c(3,20),cexRow = 1.5,cexCol = 1.2)
#tiff(file = '/Users/ll/Desktop/work/analysis/analysis_on_expression/figures/Camera gene test result (reorder).tiff')
```
    
## The interactive plot (zoom and pan)

```{r interactive,echo = FALSE,dpi = 200}
heatmaply(toheatmap.2,main = 'Camera gene test result (ordered) for Goslim set',Rowv = FALSE,dendrogram = 'col',breaks = c(-1,-0.95,0,0.95,1),col = my_palette,trace="none",row_side_colors = uc_group.2,density.info = 'density',key.title = '1 - P-value for \ncamera gene set test result',keysize = 0.3,cexRow = 0.9,cexCol = 0.6,colorbar_xanchor = 'left',colorbar_yanchor = 'bottom',margins = c(2,210,30,2))
```

        
##Hallmark of genes

```{r hallmark, include=FALSE}
h.all<-getGmt('~/Desktop/work/analysis/analysis_on_expression/gset/h.all.v6.2.symbols.gmt')
h.all.id<-geneIds(h.all)
ids<-ids2indices(h.all.id,gene.name)
#result.camera<-lapply(genes.toanalyze,camera.targetset)
#names(result.camera)<-genes.toanalyze
#save(result.camera,file = '/Users/ll/Desktop/work/analysis/analysis_on_expression/hallmark_camera.Rdata')
```

```{r heatmap2,echo = FALSE,fig.width= 120, fig.height= 80}
load('/Users/ll/Desktop/work/analysis/analysis_on_expression/hallmark_camera.Rdata')
hallmark.name<-names(h.all)
tomatch <- hallmark.name
camera.matrix<-sapply(c(1:length(result.camera)),converttopvalue)
colnames(camera.matrix)<-names(result.camera)
toheatmap.3<-data.matrix(camera.matrix)
my_palette <- colorRampPalette(c("dodgerblue3", "white", "deeppink3"))(n = 4)
heatmap.2(toheatmap.3,main = 'Camera gene test result (reorder) for common cancer hallmark set',Rowv = TRUE,dendrogram="both",breaks = c(-1,-0.95,0,0.95,1),col = my_palette,trace="none",density.info = 'density',key.title = '1 - P-value for \n camera gene set test result',keysize = 0.6,margins = c(3,19),cexRow = 1.5,cexCol = 1.2)
#tiff(file = '/Users/ll/Desktop/work/analysis/analysis_on_expression/figures/Camera gene test result (reorder) hallmark.tiff')

```

## The interactive plot (zoom and pan)

```{r interactive 2, echo = FALSE,dpi = 150}
heatmaply(toheatmap.3,main = 'Camera gene test result (reorder) for common cancer hallmark set',Rowv = TRUE,dendrogram="row",breaks = c(-1,-0.95,0,0.95,1),col = my_palette,trace="none",density.info = 'density',key.title = '1 - P-value for \n camera gene set test result',keysize = 0.6,cexRow = 0.7,cexCol = 0.6,margins = c(2,210,30,2),colorbar_xanchor = 'left',colorbar_yanchor = 'bottom')
```

## Plans
  * Interpret the plots (general, specific)
  * IC50 differential drug response
  * ssgsea correlation of gene set and regulators

```{r barcodeplot,echo=FALSE,fig.height=10,fig.width= 6,include=FALSE}
#plotfig<-function(x){
#a<-result.camera[[x]]
#index.vector <- as.logical(rep(1,nrow(a)))
#P.value<-a$PValue
#axis(side =1, at = seq(1, 69, by = 1), labels = row.names(a),las=2)
#par(mar=c(3,3,3,25))
#colorlist = c("black","red")
#axiscolor = colorlist[as.numeric(P.value > 0.05) + 1 ]
#labs = row.names(a)
#plot.title <- names(result.camera)[x]
#plt<-stripchart(P.value,vertical = TRUE,main = plot.title, col.main="red")+
#axis(side = 4, at = seq(0,1,by = 1/68)[-which(P.value<0.05)],labels = labs[-which(P.value<0.05)],las = 2,cex.axis=1)+axis(side = 4, at = seq(0,1,by = 1/68)[which(P.value<0.05)],labels = labs[which(P.value<0.05)],las = 2,cex.axis=1,col.axis = 'red')
#mtext(text=labs, side=2,at=1:length(labs),col=axiscolor,adj=.5)
#text(labels=labs, col=axiscolor, x=rep(.45,length(labs)), y=1:length(labs), srt = 0, pos = 2, xpd = TRUE)
#return(plt)
#}
#plotfig((2))
```

```{r plot2,,fig.height=100,fig.width= 12,echo = FALSE,include=FALSE}
#par(mfrow=c(5,2))
#sapply(c(1:10),plotfig)
```


