---
title: "gene set tests"
author: "Bingru Sun"
date: "2018/11/21"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE)
library(limma)
library(org.Hs.eg.db)
library(annotate)
library(GEOquery)
library(grid)
library(ggplot2)
library(GSVA)
library(GO.db)
library(GSEABase)
library(data.table)
library(gplots)
library(heatmaply)
```

```{r regulator mutation input,include = FALSE}
# input data

mutation_o<-read.delim('./data/CCLE/CCLE_DepMap_18Q1_maf_20180207 (1).txt')
mutation<-mutation_o
mutation_s<-mutation_o[which(is.na(mutation_o$ExAC_AF)),]
mutation_s<-mutation_s[which(mutation_s$isDeleterious == 'TRUE'),]
mutation_s_w<-mutation_s
mutation_s_w<-data.frame(mutation_s$Hugo_Symbol,mutation_s$Tumor_Sample_Barcode)

genelist<-as.character(unique(mutation_s_w[,1]))
rs<-function(x){
  c<-as.character(mutation_s_w[which(mutation_s_w$mutation_s.Hugo_Symbol==x),2])
  return(c)
  }
a<-lapply(genelist,rs)
names(a)<-genelist
mutatedcelllines<-a
b<-lapply(mutatedcelllines,length)
mutatedcelllines_num<-unlist(b)

regset <- getGmt('./data/Gene_regulators/targetset_nodup.gmt.txt')
reg.list<-geneIds(regset)
reg.count<-lapply(reg.list,length)
reg.count<-unlist(reg.count)
  
regulator<-names(reg.count)
target <- unique(unlist(reg.list))

reg_m<-matrix(nrow=length(regulator),ncol=length(target))
colnames(reg_m)<-target
row.names(reg_m)<-regulator


for (i in 1:nrow(reg_m)){
  x <- regulator[i]
  for (y in reg.list[[i]])
  {reg_m[x,y] <- 1}
}

reg_m[is.na(reg_m)]<-0

```

```{r ccle data input}
# input expression data

load('./data/selected_cl.Rdata')

ccle.expression<-fread('./data/CCLE/CCLE_Expression_Entrez_2012-09-29.gct',skip = 2, sep = '\t')
ccle.expression_o <- ccle.expression
ccle.expression <- as.data.frame(ccle.expression)
ccle.expression <- ccle.expression[,-c(1,2)]
ccle.anno <- ccle.expression_o[,c(1,2)]
toremove <- which(ccle.anno$Description == '')
ccle.anno <- ccle.anno[-toremove,]
ccle.expression<-ccle.expression[-toremove,]


sample.info<-fread('./data/CCLE/CCLE_sample_info_file_2012-10-18.txt', sep = '\t')
cl.pos<-match(colnames(ccle.expression),sample.info$`CCLE name`)
cl.info <- sample.info[cl.pos,]
cl.toremove <- which(cl.info$`Site Primary` == 'haematopoietic_and_lymphoid_tissue')

ccle.exp <- ccle.expression[,cl.tokeep]

cl.sample <- colnames(ccle.exp)
n <- length(cl.sample)
cl<-rep(0,n)

regulators.cl <- intersect(regulator, names(mutatedcelllines))

mutation_df <- sapply(regulators.cl,function(x){return( cl.sample %in% mutatedcelllines[[x]])})

row.names(mutation_df) <- cl.sample

mutation_df <- t(mutation_df)

mt.count <- rowSums(mutation_df)
```

```{r genename}
target.genename<-ccle.anno$Description[which(ccle.anno$Description %in% target)]
reg_m_n_anno<-reg_m[,target.genename]
select.target<-function(x){return(ccle.anno$Description %in% target.genename[as.logical(reg_m_n_anno[x,])])}
target.matrix<-sapply(regulators.cl,select.target)
target.matrix<-as.data.frame(target.matrix)
colnames(target.matrix)<-regulators.cl
```

```{r fry method}
library(statmod)
fry.targetset<-function(x){
gene<-x
idx <- target.matrix[,gene]
n<- length(cl.sample)
cl.design <- matrix(nrow=n,ncol=2)
cl.design[,1] <- as.numeric(as.logical(mutation_df[gene,]))
cl.design.2<-cl.design[,1]
cl.design.2[which(cl.design[,1] == 1)]<- 0
cl.design.2[which(cl.design[,1] == 0)]<- 1
cl.design[,2]<-cl.design.2
colnames(cl.design)<-c('mutated','non_mutated')
r1 <- fry(ccle.exp, idx, cl.design,contrast=makeContrasts(mutated - non_mutated, levels=cl.design),adjust.method = 'BH')
r1$Direction <- as.character(r1$Direction)
return(as.character(r1))}
```

```{r barcodeplot,include=FALSE}
plotbarcode<-function(gene){
n<-length(cl.sample)
cl.design <- matrix(nrow=n,ncol=2)
cl.design[,1] <- as.numeric(as.logical(mutation_df[gene,]))
cl.design.2<-cl.design[,1]
cl.design.2[which(cl.design[,1] == 1)]<- 0
cl.design.2[which(cl.design[,1] == 0)]<- 1
cl.design[,2]<-cl.design.2
colnames(cl.design)<-c('mutated','non_mutated')
fit <- lmFit(ccle.exp, cl.design)
contrasts <- makeContrasts(mutated - non_mutated, levels=cl.design)
fit2 <- contrasts.fit(fit, contrasts)
fit2 <- eBayes(fit2)
fit2$genes <- ccle.anno$Description
testResults <- topTable(fit2, number=nrow(fit2))
testResults <- testResults[complete.cases(testResults),]
mylist<-target[as.logical(reg_m[gene,])]
mylist<-mylist[!is.na(mylist)]
result.test<-testResults[which(testResults$ID %in% mylist),]
result.test$Description <- rep(gene,nrow(result.test))
index.vector = fit2$genes %in% mylist
sum(index.vector)
a<-barcodeplot(fit2$coeff[,1],index = index.vector,quantiles = c(-1,1),weights.label = 'logFC')
return(a+title(gene))
}

```

```{r fry result,eval=FALSE}
regulators.mut <- regulators.cl[mt.count > 0]
regulator_fry_df <- sapply(regulators.mut,fry.targetset)
row.names(regulator_fry_df) <- c('NGenes','Direction','Pvalue','Pvalue.Mixed')
regulator_fry_df <- t(regulator_fry_df)
regulator_fry_df_o <- regulator_fry_df
regulator_fry_df <- as.data.frame(regulator_fry_df)
regulator_fry_df$Gene <- row.names(regulator_fry_df)
regulator_fry_tb <- data.table(regulator_fry_df,stringsAsFactors = FALSE)
regulator_fry_tb[,Pvalue:=as.numeric(as.character(Pvalue))]
regulator_fry_tb[,Direction:=as.character(Direction)]
save(regulator_fry_tb,file = '../data/dry_regulator_gene_set.Rdata')

```
```{r load fry}
load('./data/dry_regulator_gene_set.Rdata')
```

```{r load regulator}
load('./data/regulators_5_cl_20_tar.Rdata')
```

## Results of ROAST gene set tests of regulators which mutated in at least 5 cell lines with at least 20 targets


```{r fry result 1, echo=FALSE}
tb1 <- regulator_fry_tb[Pvalue <= 0.05 & Gene %in% selected.regulator ]
knitr::kable(tb1)
```
```{r analysis on the regulators}

regulator.class <- fread('./data/Gene_regulators/Regulator classification.txt')

gene.age <- fread('./data/Gene_regulators/geneIDs_entrez_final_phylostrata_phy1_phy2_phy3_no_TCGA.txt')

regulator.mutation <- fread('./data/Gene_regulators/Recurrent point mutations.txt')

regulator.selected <- regulator_fry_tb[Pvalue <= 0.05 & Gene %in% selected.regulator ]

```

## Barcode plots of regulators which mutated in at least 5 cell lines with at least 20 targets
```{r plot,fig.width=12, fig.height=12,echo=FALSE}
#pdf("C:/Users/bingru sun/Documents/figure.pdf",width = 12,height = 12)
par(mfrow=c(6,3))
for (gene in regulator.selected$Gene){
  plotbarcode(gene)
}
#dev.off()
```

## Results in lung cancer cell lines
```{r fry lung}
lung.cl<-cl.sample[grep('LUNG',cl.sample)]
mutation_df.lung <- mutation_df[,lung.cl]
fry.targetset.lung<-function(x){
gene<-x
idx <- target.matrix[,gene]
n<- length(lung.cl)
cl.design <- matrix(nrow=n,ncol=2)
cl.design[,1] <- as.numeric(as.logical(mutation_df.lung[gene,]))
cl.design.2<-cl.design[,1]
cl.design.2[which(cl.design[,1] == 1)]<- 0
cl.design.2[which(cl.design[,1] == 0)]<- 1
cl.design[,2]<-cl.design.2
colnames(cl.design)<-c('mutated','non_mutated')
r1 <- fry(ccle.exp[,lung.cl], idx, cl.design,contrast=makeContrasts(mutated - non_mutated, levels=cl.design),adjust.method = 'BH')
r1$Direction <- as.character(r1$Direction)
return(as.character(r1))}

lung.count <- rowSums(mutation_df.lung)

lung.mut <- regulators.cl[lung.count > 0]

regulator_fry_df.lung <- sapply(lung.mut,fry.targetset.lung)

row.names(regulator_fry_df.lung) <- c('NGenes','Direction','Pvalue','Pvalue.Mixed')

regulator_fry_df.lung <- t(regulator_fry_df.lung)

regulator_fry_df.lung_o <- regulator_fry_df.lung

regulator_fry_df.lung <- as.data.frame(regulator_fry_df.lung)

regulator_fry_df.lung$Gene <- row.names(regulator_fry_df.lung)

regulator_fry_tb.lung <- data.table(regulator_fry_df.lung, stringsAsFactors = FALSE)

regulator_fry_tb.lung[,Pvalue := as.numeric(as.character(Pvalue))]

regulator_fry_tb.lung[,NGenes := as.numeric(as.character(NGenes))]

#save(regulator_fry_df.lung,file = '../data/fry_regulator_gene_set_lung.Rdata')
```

```{r regulator analysis}
regulator.selected <- regulator_fry_tb.lung[Pvalue <= 0.05 & Gene %in% selected.regulator]

knitr::kable(regulator.selected)
```

## barcode plots of results in lung cancer cell lines
```{r plot2,fig.width=10, fig.height=20,echo=FALSE}
par(mfrow=c(5,3))
for (gene in regulator.selected$Gene){
  plotbarcode(gene)
}
```

## Results in skin cancer cell lines
```{r skin}
skin.cl<-cl.sample[grep('SKIN',cl.sample)]
mutation_df.skin <- mutation_df[,skin.cl]
fry.targetset.skin<-function(x){
gene<-x
idx <- target.matrix[,gene]
n<- length(skin.cl)
cl.design <- matrix(nrow=n,ncol=2)
cl.design[,1] <- as.numeric(as.logical(mutation_df.skin[gene,]))
cl.design.2<-cl.design[,1]
cl.design.2[which(cl.design[,1] == 1)]<- 0
cl.design.2[which(cl.design[,1] == 0)]<- 1
cl.design[,2]<-cl.design.2
colnames(cl.design)<-c('mutated','non_mutated')
r1 <- fry(ccle.exp[,skin.cl], idx, cl.design,contrast=makeContrasts(mutated - non_mutated, levels=cl.design),adjust.method = 'BH')
r1$Direction <- as.character(r1$Direction)
return(as.character(r1))}

skin.count <- rowSums(mutation_df.skin)

skin.mut <- regulators.cl[skin.count > 0]

regulator_fry_df.skin <- sapply(skin.mut,fry.targetset.skin)

row.names(regulator_fry_df.skin) <- c('NGenes','Direction','Pvalue','Pvalue.Mixed')

regulator_fry_df.skin <- t(regulator_fry_df.skin)

regulator_fry_df.skin_o <- regulator_fry_df.skin

regulator_fry_df.skin <- as.data.frame(regulator_fry_df.skin)

regulator_fry_df.skin$Gene <- row.names(regulator_fry_df.skin)

regulator_fry_tb.skin <- data.table(regulator_fry_df.skin,stringsAsFactors = FALSE)

regulator_fry_tb.skin[,Pvalue := as.numeric(as.character(Pvalue))]

regulator_fry_tb.skin[,NGenes := as.numeric(as.character(NGenes))]

#save(regulator_fry_df.skin,file = '../data/fry_regulator_gene_set_skin.Rdata')
```

```{r regulator analysis skin}
regulator.selected <- regulator_fry_tb.skin[Pvalue <= 0.05 & Gene %in% selected.regulator]

knitr::kable(regulator.selected)
```
## barcode plots of results in skin cancer cell lines

```{r plot3,fig.width=10, fig.height=10,echo=FALSE}
for (gene in regulator.selected$Gene){
  plotbarcode(gene)
}
```

```{r idx, include = FALSE}
slim_terms <- getOBOCollection("./data/Gene_regulators/goslim_generic.obo")
xx <- as.list(org.Hs.egGO2ALLEGS)
goids <- xx[2:3]
gonames<-Term(GOTERM)
go_bp_children <- GOBPOFFSPRING$"GO:0008150"
slim_terms <- ids(slim_terms)[which(ids(slim_terms) %in% go_bp_children)]
slim_terms <- cbind(slim_terms, unname(gonames[match(slim_terms, names(gonames))]))
pathways <- xx[slim_terms[,1]]
pathways <- pathways[!is.na(names(pathways))]
names(pathways) <- unname(gonames[match(names(pathways), names(gonames))])
gene.ORF <- unlist(strsplit(ccle.anno$Name, "_"))
gene.ORF <- gene.ORF[gene.ORF != 'at']
#gogsva<-gsva(e,pathways,method=c('ssgsea'))
ids<-ids2indices(pathways,gene.ORF)
```

```{r camera , include = FALSE}
camera.targetset<-function(x){
gene<-x
n<-length(cl.sample)
cl.design <- matrix(nrow=n,ncol=2)
cl.design[,1] <- as.numeric(as.logical(mutation_df[gene,]))
cl.design.2<-cl.design[,1]
cl.design.2[which(cl.design[,1] == 1)]<- 0
cl.design.2[which(cl.design[,1] == 0)]<- 1
cl.design[,2]<-cl.design.2
colnames(cl.design)<-c('mutated','non_mutated')
r1 <- camera(ccle.exp, ids, cl.design,contrast=makeContrasts(mutated - non_mutated, levels=cl.design))
#Pvalue<-r1$Pvalue
#Direction<-r1$Direction 
#Pvalue<-1-Pvalue
#Pvalue[which(Direction == 'Down')]<- -Pvalue[which(Direction == 'Down')]
return(r1)}
```


```{r regulators all}

result.camera<-lapply(selected.regulator,camera.targetset)
names(result.camera)<-selected.regulator
```

```{r process age old}
process.name<-names(pathways)
geneage<-read.csv('./data/Gene_regulators/geneIDs_entrez_final_phylostrata_phy1_phy2_phy3_no_TCGA.txt',header = TRUE,row.names = 1)
mode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}
convertgeneage<-function(x){
  process.gene<-pathways[[x]]
  process.age<-geneage[with(geneage,geneage$Entrez %in% process.gene),2]
  return(mode(process.age))
}
processage<-sapply(c(1:69),convertgeneage)
names(processage)<-process.name
#save(processage,file = 'goslim_age.Rdata')
uc_group<-ifelse(processage<4,'uc','mc')
uc_color<-ifelse(processage<4,'pink','blue')
process.name.sort<-names(sort(uc_color))
tomatch<-process.name.sort
```

```{r pvalue2}

process.name.sort<-names(sort(uc_color))
tomatch<-process.name.sort

converttopvalue<-function(x){
#  a<-result.camera[[x]]
a<-result.camera[[x]]
Pvalue<-1-a$PValue
Direction<-a$Direction
Pvalue[ifelse(Direction == 'Down',TRUE,FALSE)]<--Pvalue[ifelse(Direction == 'Down',TRUE,FALSE)]
position <- match(tomatch,row.names(a))
Pvalue.new<-Pvalue[position]
names(Pvalue.new)<-tomatch
return(Pvalue.new)
}

camera.matrix<-sapply(c(1:length(result.camera)),converttopvalue)
colnames(camera.matrix)<-names(result.camera)
toheatmap<-data.matrix(camera.matrix)
my_palette <- colorRampPalette(c("purple", 'white',"orange"))(n = 4)
```

## Camera gene set tests of GO slim biological processes under regulator mutations


```{r camera heatmap 2,fig.width = 12,fig.height = 12}
#pdf('camera_reorder.pdf',width = 15,height = 15)
hp.2 <- heatmap.2(toheatmap,main = 'Camera analysis for GO slim processes',
          #Rowv = FALSE,
          dendrogram="both",
          breaks = c(-1,-0.95,0,0.95,1),
          col = my_palette,trace="none",
          RowSideColors = sort(uc_color),
          #density.info = 'density',
          #key.title = '1 - P-value for \ncamera gene set test result',
          key = FALSE,
          #seriate = 'mean',
          #row_dend_left = TRUE,
          #plot_method = 'plotly'
          margins = c(15,28),
          cexRow = 1.2,
          cexCol = 0.8
          )
#mtext('MC',side = 2,line = -13,at = 0.7)
#mtext('UC',side = 2,line = -13,at = 0.25)
mtext('GO slim class',side = 2,line = 1)
#dev.off()
```

```{r order}
order.2 <- row.names(toheatmap)[rev(hp.2$rowInd)]
```

#### order of regulator classes
```{r regulator reg 1}
load('./data/regulator_regulation.Rdata')
reg.class <- regulator_tb$class
names(reg.class) <- regulator_tb$regulator
reg.col <- as.factor(reg.class)
reg.level <- levels(reg.col) 
levels(reg.col) <- c('lightcyan','steelblue','hotpink','white','springgreen','tan')
reg.level[order(levels(reg.col))]
reg.col <- as.character(reg.col[selected.regulator])
names(reg.col) <- selected.regulator
selected.regulator.order <- selected.regulator[order(reg.col)]
```

### Labeled camera results
```{r heatmap with column label,fig.width = 12,fig.height = 12}
#pdf('camera_label.pdf',width = 15,height = 15)
heatmap.2(toheatmap[,selected.regulator.order],main = 'Camera analysis for goslim processes',
          #Rowv = FALSE,
          dendrogram="col",
          breaks = c(-1,-0.95,0,0.95,1),
          col = my_palette,trace="none",
          RowSideColors = sort(uc_color),
          density.info = 'density',
          key.title = '1 - P-value for \ncamera gene set test result',
          keysize = 0.6,
          #seriate = 'mean',
          #row_dend_left = TRUE,
          #plot_method = 'plotly'
          margins = c(15,30),
          cexRow = 1.2,
          cexCol = 0.8,
          Colv = FALSE,
          ColSideColors= reg.col[order(reg.col)]
          )
mtext('MC',side = 2,line = -2,at = 0.85)
mtext('UC',side = 2,line = -2,at = 0.45)
mtext('goslim class',side = 2,line = 1)
#dev.off()
```

### Camera results of significant regulators in ROAST analysis and CNA analysis
```{r,fig.width=12,fig.height=12}
heatmap.2(toheatmap[,selected.regulator[regulator_tb$class == 'CNV,target set']],main = 'Camera analysis for goslim processes',
          #Rowv = FALSE,
          dendrogram="col",
          breaks = c(-1,-0.95,0,0.95,1),
          col = my_palette,trace="none",
          RowSideColors = sort(uc_color),
          density.info = 'density',
          key.title = '1 - P-value for \ncamera gene set test result',
          keysize = 0.6,
          #seriate = 'mean',
          #row_dend_left = TRUE,
          #plot_method = 'plotly'
          margins = c(15,30),
          cexRow = 1.2,
          cexCol = 0.8
          #Colv = FALSE,
          #ColSideColors= reg.col[order(reg.col)]
          )
mtext('MC',side = 2,line = -2,at = 0.85)
mtext('UC',side = 2,line = -2,at = 0.45)
mtext('goslim class',side = 2,line = 1)
```

### regulators colored with phylostrata (red is late metazoan, white is unicellular, pink is early metazoan)

```{r regulator reg 2}
reg.age <- regulator_tb$phylostrata
names(reg.age) <- regulator_tb$regulator
reg.age[reg.age %in% c(2:3)] <- 1
reg.age[reg.age %in% c(2:8)] <- 2
reg.age[reg.age %in% c(9:12)] <- 3
reg.col <- as.factor(reg.age)
reg.level <- levels(reg.col) 
colfun <- colorRampPalette((c('white','red')))
levels(reg.col) <- colfun(3)
reg.col <- as.character(reg.col[selected.regulator])
names(reg.col) <- selected.regulator
selected.regulator.order <- selected.regulator[order(reg.col)]
```

```{r another hatmap,fig.width = 12,fig.height = 12}
heatmap.2(toheatmap[,selected.regulator.order],main = 'Camera analysis for goslim processes',
          #Rowv = FALSE,
          dendrogram="col",
          breaks = c(-1,-0.95,0,0.95,1),
          col = my_palette,trace="none",
          RowSideColors = sort(uc_color),
          density.info = 'density',
          key.title = '1 - P-value for \ncamera gene set test result',
          keysize = 0.6,
          #seriate = 'mean',
          #row_dend_left = TRUE,
          #plot_method = 'plotly'
          margins = c(15,30),
          cexRow = 1.2,
          cexCol = 0.8,
          #Colv = FALSE,
          ColSideColors= reg.col[order(reg.col)]
          )
mtext('MC',side = 2,line = -2,at = 0.85)
mtext('UC',side = 2,line = -2,at = 0.45)
mtext('goslim class',side = 2,line = 1)
```


```{r save result,eval = FALSE}
regulator.fry <- regulator_fry_tb[Pvalue <= 0.05 & Gene %in% selected.regulator]
regulator.fry.lung <- regulator_fry_tb.lung[Pvalue <= 0.05 & Gene %in% selected.regulator]
regulator.fry.skin <- regulator_fry_tb.lung[Pvalue <= 0.05 & Gene %in% selected.regulator]
```

```{r save result 2,eval = FALSE}
regulator.fry <- regulator.fry$Gene
regulator.fry.lung <- regulator.fry.lung$Gene
regulator.fry.skin <- regulator.fry.skin$Gene
save(regulator.fry,regulator.fry.lung,regulator.fry.skin,file = '../data/fry_result.Rdata')
```


